---
title: "Week 5 - Introduction to Data Analysis"
author: "ETC1010-5510"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

# <span style="color:blue"> Exercise 2: Summarise and Tabulate Missing Values</span>

#### **Estimated time: 20 min**

```{r library}
library(naniar)
library(dplyr)
```

## <span style="color:green"> Section A: Summarizing missingness </span>

### Instructions: Section A

Now that you understand the behavior of missing values in R, and how to count them, let's scale up our summaries for cases (rows) and variables, using `miss_var_summary()` and `miss_case_summary()`, and also explore how they can be applied for groups in a dataframe, using the `group_by` function from `dplyr`.

Calculate summaries of missingness in the `airquality` dataset for variables using the `miss_var_summary()` function.

## Answer the following questions:

### Q1A. Summarise missingness in _each variable_ of the `airquality` dataset. Pass `airquality` into the `miss_var_summary()` function.

```{r q1-a}
# Summarise missingness in each variable of the `airquality` dataset
miss_var_summary(airquality)
```

### Q2A. Calculate summaries of missingness in the `airquality` dataset for the _cases_ using the `miss_case_summary()` function.

```{r q2-a}
# Summarise missingness in each case of the `airquality` dataset
miss_case_summary(airquality)
```



### Q3A. Using the `airquality` dataset, use `group_by()` to create summaries for _each variable_, by each `Month`

* Pass `Month` into the `group_by()` function.
* Then use the `miss_var_summary()` and `miss_case_summary()` function.

```{r q3-a}
airquality %>%
  group_by(Month) %>%
  miss_var_summary()
```

### Q4A. Using the `airquality` dataset, use `group_by()` to create summaries for _each case_, by each `Month`. 

```{r q4-a}
airquality %>%
  group_by(Month) %>%
  miss_case_summary()
```

## <span style="color:green"> Section B: Tabulate Missingness </span>

The summaries of missingness we just calculated give us the number and percentage of missing observations for the cases and variables. 

Another way to summarise missingness is by tabulating the number of times that there are 0, 1, 2, 3, missings in a variable, or in a case. 

In this exercise we are going to tabulate the number of missings in each case and variable using `miss_var_table()` and `miss_case_table()`, and also combine these summaries with the the `group_by` operator from `dplyr`. to explore the summaries over a grouping variable in the dataset.

### Instructions: Section B
For the `airquality` dataset:

- Tabulate missingness for each variable using `miss_var_table()`.
- Tabulate missingness for every case using `miss_case_table()`.
- Combine previous tabulations with `dplyr`'s `group_by()` function to create tabulations for each variable and case, by each `Month`.

## Answer the following questions:

Don't forget to use `miss_case_table(airquality)`.

Also, don't forget to group by `Month`, using `group_by(Month)` like this: `airquality %>% group_by(Month) %>% miss_case_table()`.

### Q1B. Tabulate missingness in each _variable_ and _case_ of the `airquality` dataset

```{r q1-b}

# variables
miss_var_table(airquality)

# cases
miss_case_table(airquality)
```

### Q2B. Tabulate the missingness in each _variable_, grouped by Month, in the `airquality` dataset
```{r q2-b}
airquality %>%
  group_by(Month) %>%
  miss_var_table()
```

### Q3B. Tabulate the missingness in each _case_, grouped by Month, in the `airquality` dataset
```{r q3-b}

airquality %>%
  group_by(Month) %>%
  miss_case_table()
```
