---
title: "Week 5 - Introduction to Data Analysis"
author: "ETC1010-5510"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# <span style="color:blue"> Exercise 3: Visualise Missing Values</span>

#### **Estimated time: 20 min**

```{r library}
library(naniar)
```

## <span style="color:green"> Section A: Visualising all of missingness </span>

It can be difficult to get a handle on where the missing values are in your data, and here is where visualisation can really help.

The function `vis_miss()` creates an overview visualisation of the missingness in the data. It also has options to cluster rows based on missingness, using `cluster = TRUE`; as well as options for sorting the columns, from most missing to least missing (`sort_miss = TRUE`).

### Instructions: Section A

Using the `riskfactors` dataset from `naniar`:

- Use `vis_miss()` to visualise the missingness in the data.
- Use `vis_miss()` with `cluster = TRUE` to explore some clusters of missingness.
- Use `vis_miss()` and sort the missings with `sort_miss` to arrange the columns by missingness.
- Use the options `cluster = TRUE` and `sort_miss = TRUE` inside of `vis_miss` to cluster and sort missing values

## Answer the following questions:

### Q1A. Visualise all of the missingness in the `riskfactors`  dataset. Use `vis_miss`

```{r q1a}
___(riskfactors)
```

### Q2A. Visualise and cluster all of the missingness in the `riskfactors` dataset. Use option `cluster`

```{r q2a}
___(riskfactors, ___)
```

### Q3A. visualise and sort the columns by missingness in the `riskfactors` dataset. Use option `sort`

```{r q3a}
___(riskfactors, ___)
```

## <span style="color:green"> Section B: Visualising missing cases and variables </span>

To get a clear picture of the missingness across variables and cases, use `gg_miss_var()` and `gg_miss_case()`. These are the visual counterpart to `miss_var_summary()` and `miss_case_summary()`.

These can be split up into multiple plots with one for each category by choosing a variable to facet by.

### Instructions: Section B
Using the `riskfactors` dataset:

- Visualise the number of missings in cases using `gg_miss_case()`.
- Explore the number of missings in cases using `gg_miss_case()` and facet by the variable `education`.
- Visualise the number of missings in variables using `gg_miss_var()`.
- Explore the number of missings in variables using `gg_miss_var()` and facet by the variable `education`.

What do you notice in the visualisations of the whole data compared to the faceting?

## Answer the following questions:

### Q1B. Visualise the number of missings in cases using `gg_miss_case()`
Use `gg_miss_case(riskfactors)` to get started, and `gg_miss_case(riskfactors, facet = education)` to facet by education.

```{r q1b}
___(riskfactors)
```

### Q2B. Explore the number of missings in cases using `gg_miss_case()` and facet by the variable `education`.
```{r q2b}
___(riskfactors, facet = ___)
```

> Note: You might get a warning that says Factor `education` contains implicit NA, consider using `forcats::fct_explicit_na`
> This is OK - it is to do with how NA values are recorded in the data, and is not an error.

### Q3B. Visualise the number of missings in variables using `gg_miss_var()`
```{r q3b}
___(riskfactors)
```

### Q4B. Explore the number of missings in variables using `gg_miss_var()` and facet by the variable `education`.
```{r q4b}
___(riskfactors, facet = ___)
```

### Q5B. With the riskfactors dataset, explore how the missingness changes across the `marital` variable using gg_miss_fct()

```{r q5b}
___(x = ___, fct = ___)
```

## <span style="color:blue"> This is the End of Exercise 3</span>
_______________________________________________________________________________
_______________________________________________________________________________

# Extra bits if interested

Let's practice a few different ways to visualise patterns of missingness using:

* `gg_miss_upset()` to give an overall pattern of missingness.
* `gg_miss_fct()` for a dataset that has a factor of interest: marriage.
* and `gg_miss_span()` to explore the missingness in a time series dataset.

What do you notice with the missingness and the facetting in the data?

- Explore missingness pattern of the `airquality` dataset with `gg_miss_upset()`.
- Explore how the missingness changes in the `riskfactors` dataset across the `marital` variable using `gg_miss_fct()`.

Using the `pedestrian` dataset:

- Explore how the missingness of the `hourly_counts` changes over a span of 3000 (you can also try different spans from 2000-5000).
- Explore the impact of `month` by facetting by `month`, and explore how missingness changes for a span of 1000.

### Using the airquality dataset, explore the missingness pattern using `gg_miss_upset()`

You will use the arguments `var = hourly_counts`, `span_every = 1000`, and `facet = month` inside of the `gg_miss_span()` function.

```{r}
gg_miss_upset(airquality)
```

### Using the pedestrian dataset, explore how the missingness of hourly_counts changes over a span of 3000 
```{r}
gg_miss_span(pedestrian, var = hourly_counts, span_every = 3000)
```

### Using the pedestrian dataset, explore the impact of month by facetting by month and explore how missingness changes for a span of 1000
```{r}
gg_miss_span(pedestrian, var = hourly_counts, span_every = 3000, facet = month)
```
