---
title: "Exploring French Fries"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = T, eval = T)
```

# Exercise 3

1. Open french-fries.rmd and load the tidyverse package by replacing the '___' with the correct package

```{r setup}
#install.packages(c("tidyverse", "here"))
library(tidyverse)
```

2. Read in the french fries csv data file, is `french_fries` in long or wide format, how can you tell? 

```{r read-fries}
french_fries <- read_csv("data/french_fries.csv")
french_fries
```


3. The next code chunk converts the data to long form. What do you notice about each observation?

```{r long-fries}
fries_long <- french_fries %>% 
  pivot_longer(cols = potato:painty,
               names_to = "type", 
               values_to = "rating") %>%
  mutate(type = as.factor(type))
```

4. Filter french fries data using `filter()` to have:


    (a) only week 1

```{r fries-long-filter}
fries_long %>% filter(time == 1)
```


    (b) weeks 1-4 only

```{r fries-long-filter-1-4}
fries_long %>% filter(time %in% c("1", "2", "3", "4"))
```


    (c) oil type 1 (oil type is called treatment)

```{r fries-long-filter-oil}
fries_long %>% filter(treatment == 1)
```


    (d) oil types 1 and 3 but not 2

```{r fries-long-filter-oitl-2}
fries_long %>% filter(treatment != 2)
```


5. Show the following variables using `select()`:

    (a) choose time, treatment and rep

```{r fries-long-select}
fries_long %>% select(time, treatment, rep)
```


    (b) choose subject through to rating

```{r fries-long-select-colon}
fries_long %>% select(subject:rating)
```

    (c) show everything except subject (drop subject)

```{r fries-long-drop}
fries_long %>% select(-subject)
```


6. For the french fries data compute a new variable called lrating by taking a log of the rating using `mutate`  

```{r fries-long-mutate}
fries_long %>%
  mutate(lrating = log10(rating))
```


7. Use `group_by()` and `summarise()` to do the following:  

    (a) Compute the average rating by subject  

```{r fries-long-group-by}
fries_long %>% 
  group_by(subject) %>%
  summarise(rating = mean(rating, na.rm=TRUE))
```


    (b) Compute the average rancid rating per week 

```{r fries-long-filter-group-by}
fries_long %>% 
  filter(type == "rancid") %>%
  group_by(time) %>%
  summarise(rating = mean(rating, na.rm=TRUE))
```

8. Use `count()` to count the following:

    (a) the number of subjects

```{r fries-count-subject}
fries_long %>% count(subject, sort=TRUE) 

```

    (b) the number of types

```{r fries-count-type}
fries_long %>% count(type) 
```

# Exercise 4: Data analysis

1. Are the `rating`s for each `type` similar?

```{r fries-similar-ratings}
fries_long %>% 
  group_by(type) %>%
  summarise(m = mean(rating, 
                     na.rm = TRUE), 
            sd = sd(rating, 
                    na.rm = TRUE)) %>%
  arrange(-m)

ggplot(fries_long,
       aes(x = type, 
           y = rating)) +
  geom_boxplot()

```

*Note: the plot prints but you see a warning, this is because there are 9 rows that are missing a rating*


**Answer: The scales of the ratings are quite different. Mostly the chips are rated highly on potato'y, but low on grassy.** 


2. Are the replicates (`reps`) like each other?

    (a) Pivot the fries_long dataset wide to show the `rating` for `rep` 1 and `rep` 2 in their own column

```{r fries-spread}
fries_spread <- fries_long %>% 
  pivot_wider(names_from = rep, 
              values_from = rating)
  
fries_spread

```


    (b) Plot the ratings for rep 1 and 2 and add a title to the plot with a short summary of the data quality 
    

```{r fries-spread-plot, out.width= "80%"}
  ggplot(fries_spread,
         aes(x = `1`, 
             y = `2`)) + 
  geom_point() + 
  labs(title = "Data is poor quality: the replicates do not look like each other!")

```

