---
title: "Koala Bilby Exercise"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output: html_document
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = T, eval = T)
```


# Demonstrate with koala bilby data 

Here is a little data set to practice `pivot_longer`, `pivot_wider` and `separate` on.

1. First we load the library `tidyverse` to load up all of our functions.

```{r library}
library(tidyverse)
```

2. Read the data in and explore the variables

```{r koala-bilby}
kb <- read_csv("data/koala_bilby.csv")
kb
```


3.Pivot the data into long form, naming the two new variables, `label` and `count`

```{r kb-long}
kb_long <- kb %>% 
  pivot_longer(cols = -ID,
               names_to = "label", 
               values_to = "count")

kb_long

pivot_longer(
    kb, cols=c(-ID), 
    names_to=c("animal","region"), names_sep="_", values_to="count") 
```

4.Separate the labels into two new variables, `animal`, `state`

```{r kb-separate}
kb_long_sep <- kb_long %>%
  separate(col = label, 
           into = c("animal", "state"))

kb_long_sep
```

5.Pivot the long form data into wide form, where the columns are `state`s. 

```{r kb-wide-state}
kb_long_sep %>% 
  pivot_wider(names_from = state,
              values_from = count)
```

6.Convert the long form data into wide form, where the columns are the `animal`s. 

```{r kb-pivot-wider-animal}
kb_long_sep %>% 
  pivot_wider(names_from = "animal",
              values_from = "count")
```
