---
title: "FIT3152asm_1_Q2"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# no scientific notation
options(scipen = 999)
```
Import needed library
```{r import-lib, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
```
Read the data
```{r read data}
rm(list = ls())
set.seed(31084222)
data <-  read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Assignment_FIT3152_2021/webforum.csv")

data <- data[sample(nrow(data),20000),] #20000 rows
```
Clean the data
```{r min_max_normalisation }
#define Min-Max normalisation method
min_max_normalisation <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }
```

```{r data modification}
data$Date <- as.Date(data$Date)


#check if there is any missing values
sum(is.na(data))

data_tidy <- data %>%
  mutate(month = month(Date,  label = TRUE, abbr = TRUE), 
         wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1),
         year = year(Date),
         day = day(Date),
         hour =  hour(hm(data$Time)))

normalise_data_tidy <- function(x){
  #apply Min-Max normalisation to all numeric columns
data_tidy_norm <- as.data.frame(lapply(x[,5:19], min_max_normalisation))
return(data_tidy_norm)

}
data_tidy_norm <-  normalise_data_tidy(data_tidy)


#scale the data
data_tidy_scale <- as.data.frame(scale(data_tidy[5:19]))

```




==============================================================================================================
- Q2

    Analyse the language used by groups. Some starting points:
    
```{r corr-of-thread}
df_cor_thread <- data_tidy_norm %>%
  cbind(ThreadID = data_tidy$ThreadID)%>%  
  group_by(ThreadID)%>%
summarise(Analytic_m = mean(Analytic)*100,
                                             Clout_m = mean(Clout)*100,
                                             Authentic_m = mean(Authentic)*100,
                                            WC_m = mean(WC)*100,
                                            WPS_m = mean(WPS)*100,
                                            affect_m = mean(affect)*100,
                                             i_m = mean(i)*100,
          we_m = mean(we)*100,
                    they_m = mean(they)*100,
                    number_m = mean(number)*100,


                                             )


summary(df_cor_thread)
```
    
 

    b By analysing the linguistic variables for all or some of the threads, is it possible to see a
    difference in the language used by different groups?
    --------------------------------------------------------------------------------------------
    
    1. find out the 10 most active threads
    
```{r top-10-Active}
data_tidy_norm <- data_tidy_norm %>% cbind(ThreadID = data_tidy$ThreadID)
df_active_10 <-  data_tidy %>% group_by(ThreadID)%>% summarise(count = n()) %>% arrange(desc(count))
df_active_10 <- head(df_active_10,10)
df_active_10

df <-  data_tidy %>% filter(data_tidy_norm$ThreadID %in% df_active_10$ThreadID  ) %>% arrange(ThreadID)

```

```{r corrgram-top10}
library("corrgram")
corrgram(df,upper.panel=panel.cor)
```

    
    Relationship :
    2. corrogram 
    3. see if there is any relationship
    
    Tone : 
    4. Which one is the most positvie
```{r which-+ve}
data_nega_pose = df %>% group_by(ThreadID) %>% summarise(mean = mean(Tone))
data_nega_pose = data_nega_pose %>% mutate(emo = ifelse(mean >50 , "Positive", "Negative"))%>% 
  group_by(emo) %>% arrange(desc(mean))
```

    What languages used to shape this Tone?
    1.Pick the most postive and negative one to analyse 
    
```{r}

#most positve
data.mostPose <- data_tidy %>% filter(ThreadID =="472752" )
data.mostPose <- data.mostPose[5:19]
# corrgram(data.mostPose,upper.panel=panel.cor)

#most negative
data.mostNega <- data_tidy %>% filter(ThreadID =="309286" )
data.mostNega <- data.mostNega[5:19]
# corrgram(data.mostNega,upper.panel=panel.cor)



```


   
```{r radar-chart-all}
library(fmsb)
data.mostPose <- data.mostPose %>% select(Analytic,Clout,Authentic,WC,WPS,affect)%>% 
  mutate(Pronoun = data.mostPose[,7] + data.mostPose[,8]+ data.mostPose[,9] +  data.mostPose[,10] )


data.mostPose <- as.data.frame(lapply(data.mostPose[,], min_max_normalisation))
data.mostPose <- data.mostPose %>% summarise(Analytic_m = mean(Analytic)*100,
                                             Clout_m = mean(Clout)*100,
                                             Authentic_m = mean(Authentic)*100,
                                            WC_m = mean(WC)*100,
                                            WPS_m = mean(WPS)*100,
                                            affect_m = mean(affect)*100,
                                             Pronoun_m = mean(Pronoun)*100,
                                             )

# negative
data.mostNega <- data.mostNega %>% select(Analytic,Clout,Authentic,WC,WPS,affect)%>% 
  mutate(Pronoun = data.mostNega[,7] + data.mostNega[,8]+ data.mostNega[,9] +  data.mostNega[,10] )


data.mostNega <- as.data.frame(lapply(data.mostNega[,], min_max_normalisation))
data.mostNega <- data.mostNega %>% summarise(Analytic_m = mean(Analytic)*100,
                                             Clout_m = mean(Clout)*100,
                                             Authentic_m = mean(Authentic)*100,
                                            WC_m = mean(WC)*100,
                                            WPS_m = mean(WPS)*100,
                                            affect_m = mean(affect)*100,
                                             Pronoun_m = mean(Pronoun)*100,
                                             )

radar_data_Pose <- data.mostNega %>% rbind(data.mostPose )



radar_data_Pose <-data.frame( Analytic = c(100, 0 , data.mostPose[1,1],data.mostNega[1,1] ),
                              Clout = c(100, 0 , data.mostPose[1,2],data.mostNega[1,2]  ),
                              Authentic = c(100, 0 , data.mostPose[1,3],data.mostNega[1,3]  ),
                              WC = c(100, 0 , data.mostPose[1,4],data.mostNega[1,4]  ),
                              WPS = c(100, 0 , data.mostPose[1,5],data.mostNega[1,5]  ),
                              affect = c(100, 0 , data.mostPose[1,6],data.mostNega[1,6]  ),
                              Pronoun = c(100, 0 , data.mostPose[1,7],data.mostNega[1,7]  ),
                              row.names = c("max","min","P","N")
                              )
#defien the colors filled
colors_fill <- c(scales::alpha("blue", 0.3),scales::alpha("red", 0.5))
#define the line colors
colors_line <- c(scales::alpha("black", 0.5),scales::alpha("darkgrey", 0.5))
radarchart(radar_data_Pose, axistype = 1,
           seg = 2,
    # Customize the polygon
     pfcol =colors_fill, plwd = 2, plty = 1,
    # Customize the grid
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    pcol = colors_line,
    # Customize the axis
    axislabcol = "grey",
    caxislabels = c(0, 50, 100))
           
legend( x=0.6,y=1.35,legend = row.names( radar_data_Pose[3:4,] ) ,
        bty = "n",
        pch = 20,
        col=colors_fill,
        cex=1.2,
        pt.cex=3)

```
    

    c Does the language used within threads (or between threads) change over time? How
        consistent or variable is the language used within threads?
    --------------------------------------------------------------------------------------------
        
      
```{r vis-lan-complexity-by-yr}

data_tidy_norm <-  normalise_data_tidy(data_tidy)
data_tidy_norm <- data_tidy_norm %>% cbind(ThreadID = data_tidy$ThreadID, Date =data_tidy$Date )


data_tidy_norm <- data_tidy_norm %>% select(Analytic,Clout,Tone,Authentic,WC,WPS,affect,ThreadID,Date, i ,we,they, you )%>% 
  mutate(LangStructure = data_tidy_norm[,1] + data_tidy_norm[,6] )

grp_yr = data.frame(data_tidy_norm ) %>% cbind(year = data_tidy$year, month=data_tidy$month )

grp_yr <- grp_yr %>% group_by(ThreadID,year,month) %>% summarise(count =n(),LangStructure = mean(LangStructure,na.rm = TRUE),)%>% 
  arrange(desc(year)) 

# visualise it
mycol <- c("navy", "blue", "cyan", "lightcyan", "yellow", "red", "red4")
ggplot(grp_yr, aes(x= year , y=month, color =LangStructure )) +
  geom_tile (aes(fill=LangStructure),colour = "white" )+labs(
    title = "Time-Series Calendar Heatmap of the language struture",
       x = "Year",
       y = "Month"
    )+ 
  scale_fill_gradientn(colours = mycol)+theme_bw()+
  theme(axis.text.x = element_text(, color="BLACK", angle=90))
```

```{r vis-pronoun-used-by-yr}

data_tidy_norm <- data_tidy_norm %>% select(ThreadID,Date, i ,we,they, you ) 


grp_yr <- data.frame(data_tidy_norm ) %>% cbind(year = data_tidy$year, month=data_tidy$month )

grp_yr <- grp_yr %>% group_by(ThreadID,year,month) %>% summarise(count =n(),i=median(i),
            we=median(we),
            you=median(you),
            they = median(they))%>% 
  arrange(desc(year)) 

grp_yr <-grp_yr %>%   pivot_longer(cols = c(`i`,`we`,`you`,`they`), names_to = "Pronoun", values_to= "Values")

mycol <- c("navy", "blue", "cyan", "lightcyan", "yellow", "red", "red4")
ggplot(grp_yr, aes(x= year , y=month, color = Values)) + 
  geom_tile (aes(fill=Values),colour = "white" ) +  
  scale_fill_gradientn(colours = mycol)+facet_grid(~Pronoun)+theme_linedraw()+
  theme(axis.text.x = element_text(, color="BLACK", angle=90))+labs(
    title = "Time-Series Calendar Heatmap: Pronoun of I, They, We, You",
       x = "Year",
       y = "Month"
    )

```

How consistent or variable is the language used within threads?
variance?
```{r consistency}


```

        
        
        
        


```{r}

data_tidy_norm <- as.data.frame(lapply(data_tidy[,5:19], min_max_normalisation))
data_tidy_norm <- data_tidy_norm %>% cbind(ThreadID = data_tidy$ThreadID, Date =data_tidy$Date )

over_df =  data.frame(data_tidy_norm )

over_df = over_df%>% group_by(ThreadID, Date) %>% summarise( WC = mean(WC,na.rm = TRUE),
            Analytic = mean(Analytic,na.rm = TRUE),
            Clout = mean(Clout,na.rm = TRUE),
            Authentic = mean(Authentic,na.rm = TRUE),
            WP = mean(WPS,na.rm = TRUE),
            i = mean(i,na.rm = TRUE),
            we = mean(we,na.rm = TRUE),
            you = mean(you,na.rm = TRUE),
            they = mean(they,na.rm = TRUE),
            number = mean(number,na.rm = TRUE),
            affect = mean(affect,na.rm = TRUE),
            posemo = mean(posemo,na.rm = TRUE),
            negemo = mean(negemo,na.rm = TRUE),
            anx = mean(anx,na.rm = TRUE))

ggplot(over_df, aes(Date,Clout)) + geom_line()+ylim(0,2)+geom_smooth()
ggplot(over_df, aes(Date,we)) + geom_line()+ylim(0,2)+geom_smooth()

```
        
        
        
        
        
        

        
        
        
        
        