---
title: "FIT3152_week2_lab"
output: html_document
---
Resource : https://rpubs.com/mcharif22/mpg-exploration
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r needed_lib, echo = TRUE}

library(tidyverse)
library(lattice)
library(ggthemes)
##library("extrafont")
```



(a) What is the relationship between city (cty) fuel consumption and highway (hwy) 
consumption? 
```{r}
## https://www.youtube.com/watch?v=qnw1xDnt_Ec
cty_VS_hwy <- 
mpg %>% 
ggplot( aes(x = hwy, y = cty, color= hwy) ) + geom_point(size=1.5, alpha = .8) + geom_smooth(method = "lm")


## add a label for x and y, title, 
cty_VS_hwy <- cty_VS_hwy+
  labs(
  title = "Relationship between ctw adn hwy",
  subtitle = "What is the relationship between city (cty) fuel consumption and highway (hwy) 
consumption?", 
x = "hwy",
y = "cty"
)


##using ggtheme to beautify the graph
cty_VS_hwy <- cty_VS_hwy + theme_few()


##add back the x and y asix titles
cty_VS_hwy <- cty_VS_hwy + theme(axis.title = element_text())

##change the color palette
##cty_VS_hwy <- cty_VS_hwy + scale_color_manual(value = "#A6611A")

## change the font
##cty_VS_hwy <- cty_VS_hwy + theme(text = element_text(family = "Serif"))
cty_VS_hwy

```

They have a positive relationship.


```{r boxplot} 

g = ggplot(data = mpg, aes(x=mpg$manufacturer,  y=mpg$hwy )) + geom_boxplot()  

## change the rotation of the text
  ##1. axis.text.x = take the text of x asix
g = g + theme (axis.text.x  = element_text(angle = 90))

## reorder based on median
order = with(mpg, reorder(mpg$manufacturer , -1*hwy, median, ))

g = ggplot(data = mpg, aes(x=order,  y=mpg$hwy )) + geom_boxplot()  

g 

```

(b) How is fuel consumption (cty/hwy) related to manufacturer, transmission, class etc.?
Are there manufacturers or car types with particularly high or low fuel consumption? 







Q4
(a) 
Which variables are the best predictors of height? 
Does this vary between men and women? 
For examples, are some variables better at predicting height in one gender over the other? 

 
```{r}

B = read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Week-(2)/body.dat.csv")

#B[25] = gender
by(B, B[25], function(df) round(cor(df[1:23], df[,24]), digit = 2))


```
So, for male :Weight          for female :ShoulderWidth  



(b) Using the same approach, which variables are best for predicting weight in each gender? 



```{r}
library(reshape2)
B1  = B[, c(1:22, 24,23,25) ]  
COR=by(B, B[25], function(df) round(cor(df[1:24], df[1:24]), digit = 2))

female = melt(COR$Female)

g= ggplot(data = female, aes(x=Var1,  y =Var2, fill=value)) + geom_tile(color = "white") 

g = g + scale_fill_gradient(low = "blue" , high = "red")
g
```


```{r}
D = read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Week-(2)/Dunnhumby1-20.csv")

g = ggplot(data = D , aes (x=D$visit_delta , y=D$visit_spend )) + geom_point()
g = g + facet_grid(customer_id ~ .) +facet_wrap(~customer_id, ncol=4)
g
```



Q3 : 

```{r} 

```
  