---
title: "Tidy text analysis 4"
author: "Your Name"
output: html_document
---

```{r setup}
library(tidyverse)
library(tidytext)

```

# The books of Jane Austin

In this lab exercise we will analyse the sentiment of Austen's books. 

Below is the code to tokenise the books and add line numbers and chapters.

```{r tidy-books}
library(janeaustenr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]", 
                                           ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

Add sentiment categories to all the books using the nrc lexicon.

```{r nrc, eval = FALSE}
nrc <- get_sentiments("---")
```


- What are the most common "anger" words used in Emma?
```{r, eval = FALSE}
Emma_nrc <- tidy_books %>% 
  filter(--- == "---") %>%
  inner_join(---, by = "---") 

Emma_nrc %>% 
  filter(--- == "anger") %>%
  count(---, sort = TRUE)
```

- What are the most common "surprise" words used in Emma?

```{r, eval = FALSE}
Emma_nrc %>% 
  filter(--- == "surprise") %>%
  count(---, sort = TRUE)

```

Using another lexicon (bing, or afinn) compute the proportion of 
positive words in each of Austen's books.

- Which book is the most positive? negative?

```{r, eval = FALSE}
wordcounts <- tidy_books %>%
  group_by(---) %>%
  summarise(total_book_words = n())

tidy_books %>% 
  inner_join(get_sentiments("bing"), by = "word") %>%
  group_by(---, sentiment) %>%
  summarize(sentiment_word_count = n()) %>%
  left_join(wordcounts, by = "book")  %>%
  mutate(ratio = sentiment_word_count / ---) %>%
  group_by(---) %>%
  top_n(1, ---)
 
```