---
title: "Correlation network"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggraph)
library(igraph)
```


# Data

In this part we are creating a data frame that contains the gene expression of 10 genes under 10 different experiements. The genes are connected in a network but we do not know what the network is, we only observe the gene expression after 10 different experiments (perturbations). Our goal here is to make inferences about the network that display the genes connections.


```{r}

# Data generation
set.seed(120)
variables_df <- data.frame(matrix(rnorm(100), ncol = 10))
names(variables_df) <- paste("Gene", c(1:10), sep = "")
rownames(variables_df) <- paste("Gene", c(1:10), sep = "")

```

# Correlations to estimate network structure
We are going to use correlations to estimate the weighted adjacency matrix so that we can find the network structure between the genes. The diagonal of the weighted adjacency matrix is set to 0 because we do not want to include the relationship of each gene with itself.

```{r}
# Correlation between genei and genej for i, i = 1, ..., 10
Correlation_matrix<- cor(variables_df)
diag(Correlation_matrix) = 0
Correlation_matrix <- abs(Correlation_matrix)
head(Correlation_matrix)

```


# Creating and edge data frame to plot the network

Think carefully what are we doing here and why are we doing it.

```{r}
# Formating correlation matrix into an edge data frame
d_edges_cor <- Correlation_matrix %>% 
  as_tibble() %>%
  mutate(from = rownames(Correlation_matrix)) %>%
  gather(to, count, -from)
head(d_edges_cor)
```


# Plotting the estimated correlation network

Here we are using the packages igraph and ggraph

```{r , echo = FALSE}

set.seed(201)
d_edges_cor %>%
  
graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = count , edge_width = count), edge_color = "cyan4") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE,
                 point.padding = unit(0.15, "lines")) +
  theme_void()

```


# What happens if we change the threshold for which edges are connected?

Assume now that we only consider edges between genes that are larger than 0.35.

How does the network changes?

```{r}

set.seed(201)
d_edges_cor %>%
  dplyr::filter(count >= 0.23) %>%
graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = count , edge_width = count), edge_color = "#89a832") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE,
                 point.padding = unit(0.15, "lines")) +
  theme_void()

```


# Explore networks with different threshold values and different colors!
