---
title: "Extension-1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(broom)
library(ggResidpanel)
```

 ## Extension exercise 1: Predict co2 at another location

Using your model, built using values collected the south pole sensor, see how well it fits values from Point Barrow, Alaska. 

1. Download the data. You can use almost the same code as for SPO but check the file name at https://scrippsco2.ucsd.edu/data/atmospheric_co2/ptb. See here:

https://scrippsco2.ucsd.edu/assets/data/atmospheric/stations/merged_in_situ_and_flask/daily/daily_merge_co2_ptb.csv

```{r read-ptb}
co2_ptb <- read_csv("data/daily_merge_co2_ptb.csv",
                    col_names = c("date", 
                                  "time", 
                                  "day", 
                                  "decdate", 
                                  "n",
                                  "flg", 
                                  "co2"), 
                    skip = 69) %>%
  mutate(lat = 71.3, 
         lon = 156.6, 
         stn = "ptb") %>%
  filter(flg == 0) %>%
  mutate(date = ymd(date)) %>% 
  mutate(day0 = day - min(day))
```

```{r gg-co2-ptb}
ggplot(co2_ptb, 
       aes(x = date, y = co2)) + 
  geom_point()
```

## Model 1: `co2 ~ day0`
```{r fit-co2_ptb}
co2_ptb_fit <- lm(co2 ~ day0, data = co2_ptb)
co2_ptb_fit
co2_ptb_fit_tidy <- tidy(co2_ptb_fit)
co2_ptb_fit_tidy
```

```{r augmented-model2-ptb}
co2_ptb_model <- augment(co2_ptb_fit, co2_ptb)
co2_ptb_model
```

```{r augmented-model-ptb}
ggplot(co2_ptb_model, 
       aes(x = date, 
           y = co2)) + 
  geom_point() + 
  geom_line(aes(y = .fitted), colour = "blue")
```

```{r residuals-date}
ggplot(co2_ptb_model,
       aes(x = .fitted, y = .resid)) +
  geom_point()
```

```{r}
library(ggResidpanel)
resid_panel(co2_ptb_fit, plots = "all")
```

## Add the quadratic term
### Model 2: `co2 ~ day0 + day0sq`
```{r daysq-mod }
co2_ptb_sq <- co2_ptb %>%
  mutate(day0sq = (day0-median(day0))^2)

co2_ptb_fit2 <- lm(co2 ~ day0 + day0sq , data = co2_ptb_sq)

tidy(co2_ptb_fit2)
```

## Plot the data, and overlay the fitted model. 
```{r daysq-mod-plot}
co2_ptb_model2 <- augment(co2_ptb_fit2, co2_ptb_sq)

ggplot(co2_ptb_model2, 
       aes(x = date, y = co2)) + 
  geom_point() + 
  geom_line(aes(y = .fitted), colour="blue")
```

### Plot the residuals
```{r daysq-resid}
ggplot(co2_ptb_model2, 
       aes(x = .fitted, y = .resid)) + 
  geom_point()
```

```{r}
library(ggResidpanel)
resid_panel(co2_ptb_fit2, plots = "all")
```
