---
title: "asm1for3152"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
needed lib
```{r}
library(tidyverse)
```

```{r setup, include=FALSE}
rm(list = ls())
set.seed(31084222)
webforum <-  read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Assignment_FIT3152_2021/webforum.csv")

webforum <- webforum[sample(nrow(webforum),20000),] #20000 rows

## as a df
webforum <- data.frame(webforum)  
```




# Question 2: How many variables and observations do we have? 
**Hint:** Look for help  *?dim* in your R console and remember
that variables are in columns and observations in rows.
*dim()* returns the number of rows and the number of columns in the data set (in that order)

```{r dim, eval = FALSE}
dim(webforum)   # 1pt
```
So, there are 276830  rows and  303 columns.


The number of variables are `dim(crime_data)---` (1pt) and the number of rows are `dim(crime_data)---` (1pt)










# Question 6: Separate the column "year" into two columns with names  "Month" and "Year". Display the first 3 lines of the data set to show the updated data set 

```{r , eval = FALSE}
##ThreadID and Author as factor
 # webforum$ThreadID = as.factor( webforum$ThreadID) 
 # webforum$AuthorID = as.factor( webforum$AuthorID) 

webforum_new <- webforum %>%
  separate(col = Date,                        # 1pt
           into = c("Year", "Month", "Day"), "-" )   # 2pt

webforum <-  webforum %>% mutate( Date = as.Date(Date))
webforum_new <-  webforum_new %>% mutate( Year = as.numeric(webforum_new$Year))
webforum_new <-  webforum_new %>% mutate( Day = as.numeric(webforum_new$Day))
webforum_new <-  webforum_new %>% mutate( Month = as.numeric(webforum_new$Month))


  head(webforum_new)                        # 1pt
```

# Question 7: If you look at the data *webforum_new*, you will notice that the variable of "Year" is coded as character. In this section, we are going to convert the variable of "Year" to a numeric variable
```{r , eval = FALSE}
webforum_new %>%
  mutate(Year = as.numeric(Year))   # 1pt
#!Dunno!
head(webforum_new)                # 1pt
```



 
#  Question 8: Display the years in the data set. How many years are included in this  data set? 
> Remember that you can learn more about what these functions by typing: `?unique` or `?length` into the R console.

```{r crime-year-summary, eval = FALSE}
unique(webforum_new$Year)    # 1pt
# length tell us the length or longitude of a variable or a vector
length(unique(webforum_new$Year))   #1pt
```



# Question 10: How many incidents do we have per "Offence_category" in total for 2019? 

```{r , eval = FALSE}
webforum_new %>% 
  dplyr::filter(Year == c("2002","2003","2004","2005","2006","2007","2008","2009","2010","2011")) %>%    # 1pt
  count(AuthorID)              # 1pt
```



The trend of the activity of the users based on WC
```{r , eval = FALSE}

year_count_wc <- webforum_new %>% 
  dplyr::filter(Year == c("2002","2003","2004","2005","2006","2007","2008","2009","2010","2011")) %>%    # 1pt
  count(Year , wt = WC, sort = TRUE)   

ggplot(year_count, aes(Year,n ) )+geom_col()+theme_bw()

ggplot(year_count, aes(Year,n, group = 1 ) )+geom_line()+geom_point()

```


The trend of the activity of the users based on WC
```{r , eval = FALSE}


month_count_wc <- webforum_new %>% 
    dplyr::filter(Month == c(1:12)) %>%    # 1pt
  count(Month , wt = WC, sort = TRUE)   

ggplot(month_count_wc, aes(Month,n ) )+geom_col()+theme_bw()


```


The trend of the activity of the users based on Weekday


```{r , eval = FALSE}


weekday_count_activity <- webforum  %>% mutate(Weekday = weekdays(as.Date(webforum$Date,'%d-%m-%y')))



weekday_count_activity <- weekday_count_activity %>% 
   group_by(Weekday)%>%   count(Weekday) 

##make weekday as a factor ordered by weekday
day_count_wc$Weekday <- ordered(day_count_wc$Weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", 
"Friday", "Saturday", "Sunday")) 



ggplot(day_count_wc, aes(Weekday,n)) +geom_col() +ylim(0,5000) 
```




The trend of the activity of the users based on Month


```{r , eval = FALSE}


month_count_activity <- webforum  %>% mutate(Weekday = weekdays(as.Date(webforum$Date,'%d-%m-%y')))



month_count_activity <- webforum_new %>% 
   group_by(Month)%>%   count(Month) 


ggplot(month_count_activity, aes(Month,n)) +geom_col() +ylim(0,3000) 

```



The trend of the activity of the users based on Year

```{r , eval = FALSE}


year_count_activity <- webforum_new %>% 
   group_by(Year)%>%   count(Year) 


ggplot(year_count_activity, aes(Year,n)) +geom_col() +xlab("Year")
  
```
```{r}
day_count_wc <- day_count_wc %>% 
   group_by(Weekday)%>%   count(AuthorID )

```



Avg of Clout by year
```{r , eval = FALSE}

year_mean_Clout <- webforum_new  %>%
  group_by(Year)%>% 
summarise(mean_Clout = mean( Clout, sort = TRUE)   )
  # count(Year , wt = Clout, sort = TRUE)   

ggplot(year_mean_Clout 
, aes(Year,mean, group = 1 ) )+geom_line()+geom_point()

```

Avg of Analytic by year
```{r , eval = FALSE}

year_mean_Analytic <- webforum_new  %>%
  group_by(Year)%>% 
summarise(mean_Analytic = mean( Analytic, sort = TRUE)   )
  # count(Year , wt = Clout, sort = TRUE)   

ggplot(year_mean_Analytic 
, aes(Year,mean, group = 1 ) )+geom_line()+geom_point()

```
Median & Mean of Tone by year
```{r , eval = FALSE}
## median
year_median_Tone <- webforum_new  %>%
  group_by(Year)%>% 
summarise(median = median( Tone, sort = TRUE)   )
  # count(Year , wt = Clout, sort = TRUE)   

ggplot(year_median_Tone , aes(Year,median, group = 1 ) )+geom_line()+geom_point()


## mean
year_mean_Tone <- webforum_new  %>%
  group_by(Year)%>% 
summarise(mean_Tone = mean( Tone, sort = TRUE)   )
  # count(Year , wt = Clout, sort = TRUE)   


ggplot(year_mean_Tone , aes(Year,mean, group = 1 ) )+geom_line()+geom_point()
```
Median & Mean of Authentic by year

```{r}
year_mean_Authentic <- webforum_new  %>%
  group_by(Year)%>% 
summarise(mean_Authentic = mean( Authentic, sort = TRUE)   )
```
```{r}
year_mean <- webforum_new  %>%
  group_by(Year)%>% 
summarise(mean_WC = mean( WC, sort = TRUE)   )


year_mean <- year_mean  %>%
cbind(mean_Analytic=year_mean_Analytic$mean_Analytic)

year_mean <- year_mean%>%
cbind(mean_Clout=year_mean_Clout$mean_Clout)


year_mean <- year_mean%>%
cbind(mean_Tone=year_mean_Tone$mean_Tone)

year_mean <- year_mean%>%
cbind(mean_Authentic =year_mean_Authentic$mean_Authentic)

year_mean
```
Relationship between "I" and negative emotion

```{r}
##take the col of i and negamo 
df <- webforum %>%select(i, negemo)

    outliers <- function(x) { 
     
      Q1 <- quantile(x, probs=.25) 
      Q3 <- quantile(x, probs=.75) 
      iqr = Q3-Q1 
     
     upper_limit = Q3 + (iqr*1.5) 
     lower_limit = Q1 - (iqr*1.5) 
     
     x > upper_limit | x < lower_limit 
    } 
     
    remove_outliers <- function(df, cols = i,negemo ) { 
      for (col in cols) { 
        df <- df[!outliers(df[[col]]),] 
      } 
      df 
    } 

   df <-  remove_outliers(df, c('i', 'negemo')) 
    
    ggplot(df, aes(i, negemo))+geom_point()
ggplot(webforum, aes(i, negemo))+geom_point()


```

Relationship between "we" and negative emotion

```{r}
##take the col of i and negamo 
df <- webforum %>%select(we, negemo)

    outliers <- function(x) { 
     
      Q1 <- quantile(x, probs=.25) 
      Q3 <- quantile(x, probs=.75) 
      iqr = Q3-Q1 
     
     upper_limit = Q3 + (iqr*1.5) 
     lower_limit = Q1 - (iqr*1.5) 
     
     x > upper_limit | x < lower_limit 
    } 
     
    remove_outliers <- function(df, cols = we,negemo ) { 
      for (col in cols) { 
        df <- df[!outliers(df[[col]]),] 
      } 
      df 
    } 

   df <-  remove_outliers(df, c('we', 'negemo')) 
    
    ggplot(df, aes(we, negemo))+geom_point()
ggplot(webforum, aes(we, negemo))+geom_point()


```


Relationship between "affect" and positive emotion

```{r}
##take the col of affect and posemo 
df <- webforum %>%select(affect, posemo)

    outliers <- function(x) { 
     
      Q1 <- quantile(x, probs=.25) 
      Q3 <- quantile(x, probs=.75) 
      iqr = Q3-Q1 
     
     upper_limit = Q3 + (iqr*1.5) 
     lower_limit = Q1 - (iqr*1.5) 
     
     x > upper_limit | x < lower_limit 
    } 
     
    remove_outliers <- function(df, cols = affect,negemo ) { 
      for (col in cols) { 
        df <- df[!outliers(df[[col]]),] 
      } 
      df 
    } 

   df <-  remove_outliers(df, c('affect', 'negemo')) 
    
    ggplot(df, aes(affect, posemo))+geom_point()
ggplot(webforum, aes(affect, posemo))+geom_point()



```

```{r}
 webforum$sum = webforum$i + webforum$we + webforum$you +webforum$they + webforum$number

df <- webforum

  outliers <- function(x) { 
     
      Q1 <- quantile(x, probs=.25) 
      Q3 <- quantile(x, probs=.75) 
      iqr = Q3-Q1 
     
     upper_limit = Q3 + (iqr*1.5) 
     lower_limit = Q1 - (iqr*1.5) 
     
     x > upper_limit | x < lower_limit 
    } 
     
    remove_outliers <- function(df, cols = sum,WPS ) { 
      for (col in cols) { 
        df <- df[!outliers(df[[col]]),] 
      } 
      df 
    } 

   df <-  remove_outliers(df, c('we', 'negemo')) 
    
    ggplot(df, aes(sum, WPS))+geom_point()
ggplot(webforum, aes(sum, WPS))+geom_point()


```



```{r}


df <- webforum %>% select(WC,Analytic,Clout,Authentic,Tone,WPS,i,we,you
                          ,they,number,affect,posemo,negemo,anx)

library(corrr)

df_mt <-  View(correlate(df))


library("corrgram")
corrgram(df,upper.panel=panel.cor)
```   
The emotion is moderrately related to the tone of the author

e.g when looking at the corr between  posemo and tone is positively corr. For negamo and tone is negatively corr


```{r}
ggplot(webforum, aes(affect, posemo)) + geom_point() +ylim(0,75)
```

