---
title: "Q1_Assignment_1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
needed lib
```{r}
library(tidyverse)
library(lubridate)
```

```{r setup, include=FALSE}
rm(list = ls())
set.seed(31084222)
data <-  read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Assignment_FIT3152_2021/webforum.csv")

data <- data[sample(nrow(data),20000),] #20000 row
```

a. Analyse activity and language on the forum over time. Some starting points:
• Describe your data: 

- How active are participants, 
```{r data modification}
data$Date <- as.Date(data$Date)


#check if there is any missing values
sum(is.na(data))

data_tidy <- data %>%
  mutate(month = month(Date,  label = TRUE, abbr = TRUE), 
         wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1),
         year = year(Date),
         day = day(Date),
         hour =  hour(hm(data$Time)))

  


#scale the data
data_tidy_scale <- as.data.frame(scale(data_tidy[5:19]))

```


```{r month}

#Month
month_df = data_tidy %>% group_by(month) %>% summarize(count =n())
ggplot(month_df,
       aes(x = month, 
           y = count)) + 
  geom_col()


```

```{r wday}


#Week-Day 
wday_df = data_tidy %>% group_by(wday) %>% summarize(count=n())
ggplot(wday_df,
       aes(x = wday, 
           y = count)) + 
  geom_col()



```


```{r hour}


#Week-Day 
library(lubridate)
hour_df = data_tidy %>% group_by(hour) %>% summarize(count=n())
ggplot(hour_df,
       aes(x = hour, 
           y = count)) + 
  geom_col()+theme_classic()+geom_smooth()


```

```{r year}



#year
year_df = data_tidy %>% group_by(year) %>% summarize(count =n())
ggplot(year_df,
       aes(x = year, 
           y = count)) + 
  geom_col()


```


```{r day}


#day
day_df = data_tidy %>% group_by(day) %>% summarize(count =n())
ggplot(day_df,
       aes(x = day, 
           y = count)) + 
  geom_col()


```
```{r number of posts per day}
data_day <- data_tidy %>% 
  group_by(Date,hour) %>%
  summarise(Count = n())
data_day
data_day <- data_day %>%
  mutate(year = year(Date))
         
```

```{r the average number of posts per weekday}
data_day %>%
  mutate(wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1)) %>% 
  group_by(wday) %>%
  summarise(m = mean(n()))

```

```{r day}
library(sugrrants)
a = data_day %>% filter(year == 2006)
data_tidy_calendar <- 
  frame_calendar(a,
                 x =hour, 
                 y = Count, 
                 date = Date, 
                 )
p1 <- ggplot(data_tidy_calendar,
       aes(x = .hour, 
           y = .Count, 
           group = Date)) +
  geom_line()
prettify(p1)



```

```{r day}
a = data_day %>% filter(year == 2003)
data_tidy_calendar <- 
  frame_calendar(a,
                 x =hour, 
                 y = Count, 
                 date = Date, 
                 )
p1 <- ggplot(data_tidy_calendar,
       aes(x = .hour, 
           y = .Count, 
           group = Date)) +
  geom_line()
prettify(p1)
```

```{r day}
df <- data_tidy %>% 
  group_by(ThreadID) %>%
  summarise(pos = sum(posemo, na.rm = TRUE),
            neg = sum(negemo, na.rm = TRUE),
            anxiety = sum(anx, na.rm =TRUE))



```
How positvie the treads overall

```{r day}
ggplot(df,aes(x = ThreadID, y = pos))+ geom_point() + ylim(0,)

```
How negative the treads overall


```{r day}
ggplot(df,aes(x = ThreadID, y = neg))+ geom_point() + ylim(0,20)

```
How anxious the treads overall


```{r day}
ggplot(df,aes(x = ThreadID, y = anxiety))+ geom_point() + ylim(0,20)

```


```{r day}
df_2 <- data_tidy %>% 
  group_by(ThreadID) %>%
  summarise(Tone = median(Tone, na.rm = TRUE))

df_2 = df_2 %>% mutate(emo = ifelse(Tone >50 , "Positive", "Negative"))%>% 
  group_by(emo) %>% summarize(count =n())

ggplot(df_2, aes(x= emo, y =count))+ geom_col()



```





- are there periods where this increases or decreases? Is there a trend over time?
- Looking at the linguistic variables,
    - do these change over time?
    - Is there a relationship between variables?