---
title: "Week 4 - Workshop 2: Exploring Women's AFL - Solution"
author: "YOUR NAME"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T)
```

```{r setup-lib}
library(tidyverse)
```

# Preparation

We are looking at data for Women's AFL.

There are two tables:

- players.csv
- teams.csv 

The data was sourced from this website: https://github.com/ropenscilabs/ozwomensport/tree/master/AFLW/data

Read them into R:

```{r read-data}
# to clean up the names of the data
library(janitor)
# clean_names is this nice function that makes nice variable names
# so you don't have to refer to vars like `with spaces` --> with_spaces.
players <- read_csv("data/players.csv") %>% clean_names() 
teams <- read_csv("data/teams.csv") %>% clean_names()
```

Have a look at the data with `View()` or `glimpse()`.

```{r glimpse}
glimpse(teams)
glimpse(players)
```


# Questions

## Q1

Using the teams dataset, make a barchart of the **total kicks** by **club**.  What do you learn?

```{r barchart-total-kicks}
ggplot(teams,
       aes(x = club,
           y = kicks_tot)) + 
  geom_col()
```

## Q2

Take that same barplot and reorder the club by `kicks_tot`, what changes?

```{r barchart-total-kicks-flip}
ggplot(teams,
       aes(x = reorder(club, kicks_tot),
           y = kicks_tot)) + 
  geom_col()
```

```{r barchart-total-kicks-flip}
ggplot(teams,
       aes(x = reorder(club, kicks_tot),
           y = kicks_tot)) + 
  geom_col() + 
  coord_flip()
```

## Q3

To compare teams, should you use average or total numbers? Explain your thinking.

**Answer:**
> Average would be a better comparison, because not all teams may have played the same number of games.

## Q4

Make a scatterplot of average kicks by average handballs. Use an aspect ratio of 1. What do you learn?

```{r gg-kicks-handballs}
ggplot(teams,
       aes(kicks_avg,
           handballs_avg)) + 
  geom_point() + 
  theme(aspect.ratio = 1)
```

> Nothing can be said about the plot per se. Perhaps `plotly()` can provide more information on individual observation.

##  Q5

Use the `plotly` package to make the previous plot interactive (with the `ggplotly` function). Report the name of the team(s) that has(ve) the highest on both variables.

```{r plotly}
library(plotly)
ggplot(teams,
       aes(kicks_avg,
           handballs_avg,
           # this label option gets passed to ggplotly
           label = club)) + 
  geom_point() + 
  theme(aspect.ratio = 1)

# ggplotly uses some programming magic and uses the last ggplot object you made
# to make it interactive.
ggplotly()
```

> Collingwood has the highest kick_avg

> Melbourne has highest handballs_avg

## Q6

Make a scatterplot of average frees for (frees_for_avg) by against (frees_agst_avg), for each season. Use an aspect ratio of 1. What do you learn? 

```{r gg-avg-frees}
ggplot(teams,
       aes(x = frees_for_avg,
           y = frees_agst_avg) ) + 
  geom_point() +
  theme(aspect.ratio = 1)
```

> Nothing can be said about the plot per se. Perhaps `plotly()` can provide more information on individual observation.

## Q7

Make the plot interactive and report the name of the team that has the most and the least free_against_avg and frees_for_avg.

```{r gg-team-frees-for-against}
# use the similar ggplotly code from before
ggplot(teams,
       aes(frees_for_avg,
           frees_agst_avg,
           text = club)) + 
  geom_point() + 
  theme(aspect.ratio = 1)

ggplotly()
```

> frees against average: Collingwood has the most ; Brisbane has the least

> frees for average: Adelaide Crows has the most ; Carlton has the least

## Q8

Using the players data, make a side-by-side boxplot of the average kicks by each club. What do you learn? 

```{r gg-players-club-avg-kicks}
ggplot(players, 
       aes(club, 
           kicks_avg)) + 
  geom_boxplot()
```


## Q9

Create a ggplotly() to find out which team had the player with the highest average kicks. Do some teams appear to have "star" players, and others have more even skills in this area?

hint: This should show points for the avegare kicks for each player by club, where the ggplotly text shows the player name. 

```{r}

ggplot(players,
       aes(club,
           kicks_avg,
           text = player)) + 
  geom_point() + 
  theme(aspect.ratio = 1)

ggplotly()


```

>Erin Phillips is the star player of Adel

>Emma Kearney is the star player of WB, close to Erin Phillips
