---
title: "Week 4 - Workshop 2: Mapping flight movements"
author: "YOUR NAME"
output: html_document
---

```{r setup}
library(tidyverse)
```

```{r load-plane}
plane_N4YRAA <- read_csv("data/plane_N4YRAA.csv")
glimpse(plane_N4YRAA)
```

Load Airline travel, airport location

```{r read-airports}
airport_raw <- read_csv("data/airports.csv")
airport_raw %>%
  select(AIRPORT, 
         LATITUDE, 
         LONGITUDE, 
         AIRPORT_STATE_NAME) %>%
  glimpse()

```

The Key is the airport three letter code, 
    - this is used in the ORIGIN and DEST column of the `plane_N4YRAA` table
    - and AIRPORT in the `airport` table


First we will tidy up the `airport_raw` data and create `airport`

```{r tidy-airport}

unique(plane_N4YRAA$ORIGIN)


airport <- airport_raw %>%
  select(AIRPORT, 
         LATITUDE, 
         LONGITUDE, 
         AIRPORT_IS_LATEST, 
         DISPLAY_AIRPORT_NAME) %>%
  filter(AIRPORT_IS_LATEST == 1) %>%  # get the rows where airport_is_latest equals 1
  select(-AIRPORT_IS_LATEST)          # now hide that column because we don't need it anymore and want to clean up the dataset

airport
```
`plane_N4YRAA`, has less airports than `airports`
    - We will only keep the rows of the `airport` table, for those that appear in the `plane_N4YRAA` table by doing a left join

```{r tidy-flight}
N4YRAA_latlon <- left_join(plane_N4YRAA,
                           airport,
                           by = c("ORIGIN" = "AIRPORT")
) %>%
  rename(
    "ORIGIN_LATITUDE" = "LATITUDE",
    "ORIGIN_LONGITUDE" = "LONGITUDE"
  )

N4YRAA_latlon %>%
  select(
    ORIGIN,
    ORIGIN_LATITUDE,
    ORIGIN_LONGITUDE,
    DISPLAY_AIRPORT_NAME
  )
```

The variables ORIGIN_LATITUDE, ORIGIN_LONGITUDE, DISPLAY_AIRPORT_NAME are added to corresponding row in the plane_N4YRAA table.

Add destination coordinates

- Added spatial coordinates (lat, lon) for the origin airport
- Same needs to be done for the destination airport

Now we combine the current data with some new data to add the destinations.

```{r}
N4YRAA_latlon_dest <- left_join(
  N4YRAA_latlon,
  airport,
  by = c("DEST" = "AIRPORT" )
) %>%
  rename(
    "DEST_LATITUDE" = "LATITUDE",
    "DEST_LONGITUDE" = "LONGITUDE"
  )

N4YRAA_latlon_dest <- N4YRAA_latlon_dest %>% arrange(FL_DATE, DEP_TIME)

# Cleanup
rm(airport)
```
