---
title: "Week 4 - Workshop 2: joins - solution"
author: "YOUR NAME"
date: "08/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)
```

```{r read-in}
gap_life <- read_csv("data/gap_life.csv")
gap_life_au <- read_csv("data/gap_life_au.csv")
gap_income <- read_csv("data/gap_income.csv")
gap_income_au <- read_csv("data/gap_income_au.csv")
gap_co2 <- read_csv("data/gap_co2.csv")
gap_co2_au <- read_csv("data/gap_co2_au.csv")



```

# Section A: Understanding joins

We have information on Australian life expectancy and income:

```{r data-view}
gap_life_au
gap_income_au
```

How do we plot life expectancy against income? We need them in the same dataframe

One technique that we've looked at is `bind_cols()`, which binds dataframes together, column wise

```{r bind-data}
bind_cols(gap_life_au,
          gap_income_au)
```

But this has problems:

1. It produces messy output (country1, year1)
2. It doesn't work if the data doesn't have the same number of rows

For example, how do we add this co2 data to income or life?

```{r data-view2}
gap_co2_au
```

We can't use `bind_cols()`

```{r data-bind2, eval=FALSE}
bind_cols(gap_co2_au,
          gap_income_au)
```

We could think about a more complex approach using `filter`, and so on...

But surely this must be a problem that we encounter in data analysis?

Someone must have thought of a solution to this before?

They did! Joins!

We can use ` left_join()` to combine the income and life expectancy data

```{r}
left_join(x = gap_income_au,
          y = gap_life_au,
          by = c("country", "year"))
```

And then what about adding co2 data?

We can add on another join:

```{r}
left_join(x = gap_income_au,
          y = gap_life_au,
          by = c("country", "year")) %>% 
  left_join(gap_co2_au,
            by = c("country", "year"))
```

And we get some missing values for co2, because we don't have c02 values for 2015 and beyond.

So now we can combine that together like so:

```{r}
gap_au <- left_join(x = gap_income_au,
          y = gap_life_au,
          by = c("country", "year")) %>% 
  left_join(gap_co2_au,
            by = c("country", "year"))

gap_au
```

Now we can make a plot!

```{r}
ggplot(gap_au,
       aes(x = gdp,
           y = life_expectancy)) + 
  geom_point()
```

# Section B

## Q1

Write down why these two joins produce different results:

```{r}
left_join(gap_co2_au,
          gap_life_au)

left_join(gap_life_au,
          gap_co2_au)
```

**Answer:**

In the first line of the code chunk above, we perform a left join between gap_co2_au and gap_life_au. A left join will use gap_co2_au as the master table to which the gap_life_au records are joined. So because gap_co_2 only has 3 rows - then this limits the number of rows that gap_life_au can be joined to - only the rows in gap_life_au from the years 2012, 2013 and 2014 will be able to be joined. 

In the second line, gap_life_au is the master to which gap_co2_au will be joined. In this case, all rows from gap_life_au are used, and the three matches from gap_co2_au are joined to this for years 2012, 2013, 2014, and the rest of the unmatched fields become NA.

## Q2

What happens when we add data from New Zealand into the mix? Can you produce a dataset that has all data on gdp, life expectancy, and co2, for both Australia and New Zealand, using the joining techniques you've just learnt?

```{r}
gap <- left_join(x = gap_income,
          y = gap_life,
          by = c("country", "year")) %>% 
  left_join(gap_co2,
            by = c("country", "year"))
```

