---
title: "Week 4 - Workshop 2: Exploring Nycflights"
author: "YOUR NAME"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

```{r setup-lib}
library(tidyverse)
library(nycflights13)
```

# Section A

Look at the sample code and write down what it does

*Note: Make sure line 13 above has been run, or the flight data will not be found.* 

This code does: ____

```{r}
airports %>% filter(faa == "LGA")
```

This code does the following: 

- Step 1: Get flights with the Delta carrier 'DL', that started at the airport code 'LGA' in August
- Step 2: ______
- Step 3: ______

```{r}
delta <- flights %>%
  filter(carrier == "DL", origin == "LGA", month == 8) %>%
  left_join(airports, by = c("dest" = "faa")) %>%
  mutate(orig_lon = -73.9, orig_lat = 40.8)
```

The code below ___ 

```{r}
flgt_weath <- flights %>%
  filter(origin == "LGA") %>%
  left_join(weather, by = c("origin", "time_hour")) %>%
  filter(wind_speed > 25)

ggplot(flgt_weath, aes(x = wind_dir, y = dep_delay)) +
   geom_point(alpha=0.5) 
  
```

# Section B

Complete these exercises about the `nycflights13` data using wrangling operations, an appropriate join, and a plot. 

You can use the examples above or code from the "flight-movement.Rmd" file to provide guidance.

## Q1

Filter the flights data to contain just Delta flights ('DL'), for August

```{r}
delta <- flights %>%
  ___(___)
```

## Q2

Add the airport locations (lat, long) of the origin and destination to the flights data 

Note: Adding coordinates would be useful if you were asked to plot flights on a map, but we won't do this today

```{r}

flights_latlon <- left_join(flights,
                                   airports,
                                   by = c("origin" = "faa")) %>%
  rename("origin_latitude" = "___",
         "origin_longitude" = "___") 

flights_latlon <- left_join(___, # Now get lat and long for the destination
                            ___,
                            ___ = c("___" = "___")) %>%
  rename("___" = "___",
         "___" = "__") 
```

# Section C

Lets look at how cross winds affect airport operations. 

## Q1

Join the weather data to the flights data, you'll need to figure out which two columns to join on

```{r}

flgt_weath <- flights %>%
  left_join(weather, by = c("____", "____")) 

```


## Q2

Filter by origin airport 'EWR' and wind speeds higher than 20mph

```{r}

flgt_weath_ewr <- ___ %>%
  ___(___ == "___", ___ > ___) 

```

## Q3

Plot delay against wind direction

Does wind direction, when windspeed is stronger, affect the operations at the airport? 

```{r}
ggplot(___, 
       aes(x = ___, 
           y = ___)) +
   geom_point(alpha=0.5) 

```
