---
title: "MLR_assignment"
output: html_document
---

The model is to estimate :
  - how strong the relationship is between Clout (dependent var) and other variables (independent var)
  - the expected Clout value at certain levels of WC,Analytic,Clout,Authentic,WPS,i,we,you,they and so on
  
Assumptions of multiple linear regression : 
```{r import-lib, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(corrgram)
```
Read the data
```{r read data}
rm(list = ls())
set.seed(31084222)
data <-  read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Assignment_FIT3152_2021/webforum.csv")

data <- data[sample(nrow(data),20000),] #20000 rows
```
Clean the data

```{r lib-for-MLR, message=F}
library(Hmisc)
library(psych)
library(car)
```


```{r}
data$Date <- as.Date(data$Date)
data

#check if there is any missing values
sum(is.na(data))

data_tidy <- data %>%
  mutate(month = month(Date,  label = TRUE, abbr = TRUE), 
         wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1),
         year = year(Date),
         day = day(Date),
         hour = hm(Time))
data_tidy

#scale the numeric data
data_tidy_scale <- data.frame( scale(data_tidy[,5:19],center=TRUE,scale=TRUE) )

df <-  data_tidy_scale %>% select(WC,Analytic,Clout,Authentic,WPS,i,we,you,they,number,affect, posemo,negemo, anx)
```


```{r view-summary}
fit <- lm(affect ~ posemo +negemo, data=data_tidy)
summary(fit)
```

```{r elim-xtreme-1}
# crPlots(fit)

# Eliminate extreme values
cutoff <- 4/((nrow(data_tidy)-length(fit$coefficients)-2)) # Cook's D plot, cutoff as 4/(n-k-1)
plot(fit, which=4, cook.levels=cutoff)                        # identify D values > cutoff

plot(fit, which=5, cook.levels=cutoff)
data_tidy <- data_tidy[-which(rownames(data_tidy)    # Row names discovered in 2 rounds
    %in% c("19998", "5379", "14068")),]  
```
After refiting the model 1
```{r refit1}
fit <- lm(affect ~ posemo+negemo, data=data_tidy)
summary(fit) 
```
```{r elim-xtreme-2}
# crPlots(fit)

# Eliminate extreme values
cutoff <- 4/((nrow(data_tidy)-length(fit$coefficients)-2)) # Cook's D plot, cutoff as 4/(n-k-1)
plot(fit, which=4, cook.levels=cutoff)                        # identify D values > cutoff

data_tidy <- data_tidy[-which(rownames(data_tidy)    # Row names discovered in 2 rounds
    %in% c("19390", "16438", "5739")),]  
```
 After refiting the model 2
```{r refit2}
fit <- lm(affect ~ posemo+negemo+anx, data=data_tidy)
summary(fit)$adj.r.squared # R2=81%, F=139.5
```
```{r elim-xtreme-3}
# crPlots(fit)

# Eliminate extreme values
cutoff <- 4/((nrow(data_tidy)-length(fit$coefficients)-2)) # Cook's D plot, cutoff as 4/(n-k-1)
plot(fit, which=4, cook.levels=cutoff)                        # identify D values > cutoff

data_tidy <- data_tidy[-which(rownames(data_tidy)    # Row names discovered in 2 rounds
    %in% c("4561", "1755", "15379")),]  
```
 After refiting the model 3
```{r refit3}
fit <- lm(affect ~ posemo+negemo, data=data_tidy)
summary(fit)$adj.r.squared # R2=81%, F=139.5
```

```{r Check-multi-collinearity}

# Check for multi-collinearity with Variance Inflation Factor
# Correlated: none VIF=1, moderately 1<VIF<5, ** highly 5<VIF<10, ...
vif(fit)
```




=============================================VISUALISATION=====================================================


```{r av-plot}
avPlots(fit, main = "The partial regression on affect given posemo and negemo")
```

https://stats.stackexchange.com/questions/125561/what-does-an-added-variable-plot-partial-regression-plot-explain-in-a-multiple

https://www.youtube.com/watch?v=QX028TaLFO0

The added variable plot gives you two dimensional perspectives for any number of variables. It may be particularly revealing in higher dimensions. One often finds revealing patterns in the residuals which were not at all obvious in the observed Y. –

```{r 3d-scatterchart}
 library(scatterplot3d)
library(rgl)
plot3d(data_tidy$posemo,data_tidy$affect ,data_tidy$negemo,type = "p")
```

