3 Creating ROC Charts 

 

 The Aliens table below provides information about several aliens including a class attribute, 
IsFriendly. A decision tree has been used to classify the data and obtained confidence values 
for IsFriendly=’yes’. 

 

a) Using this information, create a ROC chart for the Alien data. 

```{r}
library(ROCR)

pConf = c(.7,.9,.4,.1,.9,.8,.3,.6,.7,.6)
pLabel = c(0,1,0,0,1,1,1,1,1,0)

cPrep = prediction(pConf,pLabel)
cPerf = performance(cPrep, "tpr","fpr")
#ROC
plot(cPerf)
abline(0,1)

#AUC
auc = performance(cPrep,"auc")
auc@y.values
```

 Remember ROC charts graph TPR vs FPR for varying confidence thresholds. You 
can use the tables below the data table to assist you with the calculations. 

 

b) What does this ROC chart tell you about the classifier? 

 

c) What is the AUC value and what does this tell you about the classifier? 

4 Creating a Lift chart using the Aliens data 

 

 If you select an alien at random, what is the chance it will be friendly?  
```{r}
clift = performance(cPrep, "lift")
plot(clift)
```

 

a. Using the data provided, what is the value of the Lift if you choose an alien from the 
 subset which the classifier is at least 80% confident of? 

 

b. Now sketch a Lift chart, using steps of 20% (i.e. 80%, 60%, 40%, 20%) 
 
  

5 Fitting a tree, cross validation and pruning in R 

 

 The Zoo data set “zoo.data.csv” contains data relating to seven classes of animals. Construct 
a decision tree using 70% training and 30% test data. Note that you will need to make sure 
the model knows that “type” is a factor. Make a table showing the actual vs predicted 
classifications  using your test data. Calculate the accuracy of your model (by calculating 
correct classifications/all classifications). Are there any classes that are particularly difficult 
to predict? 

 

Using cross validation and the “cv.tree” function see whether or not you are able to create a 
more accurate tree by pruning. Make a prediction with the new tree and compare its 
accuracy with your original tree. 

 