---
title: "Harry Potter"
author: "ETC1010-5510"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

```{r}
library(tidyverse)
library(geomnet)
```

ensure you have installed 'geomnet':

```{r}
# install.packages("remotes")
# library(remotes)
# install_github("sctyner/geomnet")
```


## Harry Potter characters

> There is a connection between two students if one provides emotional support to the other at some point in the book. 
- Code to pull the data together is provided by Sam Tyner [here](https://github.com/sctyner/geomnet/blob/master/README.Rmd#harry-potter-peer-support-network).

```{r read-hp1, echo=FALSE}
hp_chars <- read_csv("data/hp_chars.csv")
hp_edges <- read_csv("data/hp_edges.csv")
```

## We need to do a pretty gnarly full join, since we want this for each book:

```{r read-hp, echo=FALSE}
# for each book, expand the edges data and join it to characters
hp_all <- ___ %>%
  ___ %>%
  group_map(~ full_join(.x, hp_chars, by = ___)) %>%
  bind_rows(.id = "book") %>% 
  rename(___)
```

> We now have it in that form that we wanted i.e., Harry potter data as nodes and edges

```{r print-hp}
hp_all
```

## Let's plot the characters for book 1

```{r plot-hp, echo=TRUE}
hp_all %>%
  filter(book==1) %>%
ggplot(aes(from_id = from_id, 
           to_id = to_id)) + 
  geom_net(aes(colour = house, group = house, shape = gender),
           fiteach=T, 
           directed = T, 
           size = 3, 
           linewidth = .5, 
           ealpha = .5, 
           labelon = T, 
           fontsize = 3, 
           labelcolour = "black", 
           arrowsize = .5, 
           singletons = FALSE) + 
  scale_colour_manual(values = c("#941B08","#F1F31C", "#071A80", "#154C07")) + 
  theme_net() + 
  theme(panel.background = element_rect(colour = 'black'),
        legend.position="bottom")
```

## Your Turn

### Q1. Write out the 5x5 association matrix for Harry James Potter, Hermione Granger, George Weasley, Fred Weasley and Ronald Weasley in book 1.

```{r}
main <- c("Harry James Potter", "Hermione Granger", "George Weasley","Fred Weasley", "Ronald Weasley")
hp_subset <- 
```

### Q2. In the philosphers stone, who has the most connections?
> Put your data wrangling skills to use.
> Hint: Transform to long form

```{r}

```

### Q3. In the philosphers stone, who has the least connections?
```{r}

```

### Q4. Across all the books, can you show the number of connections?
```{r}
# Transform to long format
hp_long <- 
  

hp_count <- 
  
# Can you make a plot of the number of connections across the books?
ggplot
```

### Q5. Plot the characters for all books
```{r}

hp_all %>%
ggplot(aes(from_id = ___, 
           to_id = ___)) + 
  geom_net(aes(colour = ___, group = ___, shape = ___),
           fiteach=T, 
           directed = T, 
           size = 3, 
           linewidth = .5, 
           ealpha = .5, 
           labelon = T, 
           fontsize = 3, 
           labelcolour = "black", 
           arrowsize = .5, 
           singletons = FALSE) + 
  scale_colour_manual(values = c("#941B08","#F1F31C", "#071A80", "#154C07")) + 
  ___(~___, labeller = "label_both", ncol=3) + 
  theme_net() + 
  theme(panel.background = element_rect(colour = 'black'),
        legend.position="bottom")
```


## Try exploring the other datasets in the `geomnet` package:

```{r}
library(geomnet)

?geomnet::bikes
?geomnet::blood
?geomnet::email
```

