---
title: "Exercise 7B Part Two"
author: "YOUR NAME"
output: html_document
---

```{r pkgs}
library(tidyverse)
library(readxl)
library(janitor)
library(broom)
```


## Your turn: new data - fertility data

- Using the fertility data from the [gapminder web site](https://www.gapminder.org/data/) (in data/ folder)
- This contains data of number of babies born per woman
- You can find this out by exploring the data in the aforementioned gapminder website (https://www.gapminder.org/data/documentation/gd008/)
-  Conduct the same analysis as in the gapminder example from the lecture ( ida-exercise-7b.Rmd), but for life expectancy.
- Find unusual countries

The code below will help you read in the data and process it, but you will need to make some changes to do the full analysis.

```{r read-data}
fertility_raw <- read_xlsx(here::here(
  "Week7-Workshop2/ExtraPractice/data-raw/indicator-undata-total-fertility.xlsx"
  )) %>%
  clean_names() %>%
  rename(country = total_fertility_rate)

fertility_raw
```

You will need to pivot the data into longer form, using the `pivot_longer` function. 

You should know how to do the rest from your experience with past assignments.

```{r tidy-fertility}
fertility <-
  fertility_raw %>%
  pivot_longer(cols = -country, # everything but country,
               # what is the name of the new variable we are changing the
               # column of names to?
               names_to = "year_collected",
               # What is the name of the column we are changing the values to?
               values_to = "babies_per_woman") %>%
  #extract out the year information
  mutate(year = parse_number(year_collected)) %>%
  # filter so we only look at years above 1950
  filter( year > 1950) %>%
  # center year around 1950
  na.omit %>%
  mutate(year1950 = year - 1950)
```

plot the fertility data for each year

```{r eval=FALSE}
ggplot(fertility, 
       aes(x = year, 
           y = babies_per_woman, 
           group = country)) +
  geom_line(alpha=0.1)
```

Now try and answer these questions below. You can use the examples in `ida-exercise-7b.Rmd` to help you answer these questions.

# How has fertility changed over years? For each country?
A little decrease can be seen
Major dips after 1980


## How has fertility changed in Australia?
```{r}

oz <- fertility %>% filter(country == "Australia")

oz

ggplot(data = oz, 
       aes(x = year, 
           y = babies_per_woman)) + 
  geom_line()

```
Dip after 1960.

## Fit a model for Australia
```{r lm-oz-gapminder}
oz_lm <- lm(babies_per_woman ~ year, data = oz)

oz_lm

summary(oz_lm)

oz_aug <- augment(oz_lm, oz)

oz_aug


```

## Explore model fit for Australia

```{r oz-gap-aug, fig.height=3, fig.width=8}
ggplot(data = oz_aug, 
       aes(x = year, 
           y = .fitted)) + 
  geom_line(colour = "blue") + 
  geom_point(aes(x = year,
                 y = babies_per_woman))
```
In Australia, for a unit inctrease in year, the ferility has decreased by -0.02883 on average
74.91% variance is explanined by this model.

## How has fertility increased every year / decade in Australia?
for a unit inctrease in year, the ferility has decreased by -0.02883 on average.
The increase prior to 1960, isn't explained by this model.

## How well does the model fit?
```{r oz-gap-year-resid, fig.height=3, fig.width=8}
ggplot(data = oz_aug, 
             aes(x = year, 
                 y = .std.resid)) + 
  geom_hline(yintercept = 0,
             colour = "white", 
             size = 2) +
  geom_line() 

```

74.91% variance is explanined by this model.
A lot of unexplained variance.
Fertility has decreased 0.28 years every decade, on average.

## Fit a linear model for all countries
```{r}
by_country <- fertility %>% 
  select(country, year1950, babies_per_woman) %>%
  group_by(country) %>% 
  nest()

by_country

mapped_lm <-map(by_country$data, function(x){
  lm(babies_per_woman ~ year1950, data = x) 
  })

mapped_lm
```

# Map inside data
```{r map-country}
country_model <- by_country %>% 
  mutate(model = map(data, function(x){
    lm(babies_per_woman ~ year1950, data = x)
    })
    )

country_model
```
```{r map-country-short}
country_model <- by_country %>% 
  mutate(model = map(data, ~lm(babies_per_woman ~ year1950, data = .)))

country_model
```

```{r}
country_model %>%
  mutate(tidy = map(model, tidy))
```

```{r map-tidy-model}
country_coefs <- country_model %>%
  mutate(tidy = map(model, tidy)) %>%
  unnest(tidy) %>%
  select(country, term, estimate)

country_coefs
```

```{r tidy-unnested-coefs}
tidy_country_coefs <- country_coefs %>%
  pivot_wider(id_cols = c(term, country), 
              names_from =  term,
              values_from = estimate) %>%
  rename(intercept = `(Intercept)`)

tidy_country_coefs
```
```{r oz-unnested-coef, echo=TRUE}
tidy_country_coefs %>%
  filter(country == "Australia")
```
# Plot all models
```{r augmente-many-countries, fig.height=4}
country_aug <- country_model %>% 
  mutate(augmented = map(model, augment)) %>%
  unnest(augmented)

country_aug
```
## Plot the model fit against the data in two plots
```{r plot-gapminder-data, fig.height=4}
p1 <- fertility %>% 
  ggplot(aes(year, babies_per_woman, group = country)) +
    geom_line(alpha = 1/3) + ggtitle("Data")

p2 <- ggplot(country_aug) + 
  geom_line(aes(x = year1950 + 1950, 
                y = .fitted, 
                group = country), 
            alpha = 1/3) +
  xlab("year") +
  ggtitle("Fitted models")
```

```{r plot-print-gapminder, fig.height = 4}
library(gridExtra)
grid.arrange(p1, p2, ncol=2)
```

## plot the model coefficients with plotly

## Summarise what you have learned from the model coefficients

## Find those countries with the worst fit

