---
title: "FIT3152asm_1_final_markdown"
output:
  html_document: default
  word_document: default
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,include=T, message =FALSE)
```
Import needed library
```{r import-lib, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(extrafont) # for changing the font stlye of the graph
```
Read the data
```{r read data}
rm(list = ls())
set.seed(31084222)
data <-  read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Assignment_FIT3152_2021/webforum.csv")

data <- data[sample(nrow(data),20000),] #20000 rows
```
Clean the data
```{r min_max_normalisation }
#define Min-Max normalisation method
min_max_normalisation <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }
```
Change the font style
```{r}
changeFont <-  function(){
   theme_classic() + theme(text=element_text(family="Times New Roman", face="bold", size=12)) #Times New Roman, 12pt, Bold
}
```

```{r data modification}
data$Date <- as.Date(data$Date)


#check if there is any missing values
sum(is.na(data))

data_tidy <- data %>%
  mutate(month = month(Date,  label = TRUE, abbr = TRUE), 
         wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1),
         year = year(Date),
         day = day(Date),
         hour =  hour(hm(data$Time)))

  
#apply Min-Max normalisation to all numeric columns
data_tidy_norm <- as.data.frame(lapply(data_tidy[,5:19], min_max_normalisation))

#scale the data
data_tidy_scale <- as.data.frame(scale(data_tidy[5:19]))

```


==============================================================================================================


- Q1

    a. How active are participants, are there periods where this increases or decreases?

    




   
```{r}
over_df = data_tidy %>% group_by(Date) %>% summarise(count =n())
ggplot(over_df,
       aes(x = Date, 
           y = count)) +
  geom_line() +
  geom_smooth()+labs(
    title = "The number of active authors over the years",
       subtitle = "(2002-2011)",
       x = "Year",
       y = "Activity volume"
    )+changeFont()
```



    --------------------------------------------------------------------------------------------
    c Looking at the linguistic variables, 
      1 do these change over time?

      
```{r grp-variable-by-yr}

grp_yr = data.frame(data_tidy_norm ) %>% cbind(year = data_tidy$year)

grp_yr = grp_yr %>% group_by(year) %>% 
  summarise(count =n(),Tone = mean(Tone, na.rm = TRUE),
            WC = mean(WC,na.rm = TRUE),
            Analytic = mean(Analytic,na.rm = TRUE),
            Clout = mean(Clout,na.rm = TRUE),
            Authentic = mean(Authentic,na.rm = TRUE),
            WP = mean(WPS,na.rm = TRUE),
            i = mean(i,na.rm = TRUE),
            we = mean(we,na.rm = TRUE),
            you = mean(you,na.rm = TRUE),
            they = mean(they,na.rm = TRUE),
            number = mean(number,na.rm = TRUE),
            affect = mean(affect,na.rm = TRUE),
            posemo = mean(posemo,na.rm = TRUE),
            negemo = mean(negemo,na.rm = TRUE),
            anx = mean(anx,na.rm = TRUE))  %>% 
  arrange(desc(count))
```

```{r hour}

library(extrafont)
#Week-Day 
library(lubridate)
hour_df = data_tidy %>% group_by(hour) %>% summarize(count=n())
ggplot(hour_df,
       aes(x = hour, 
           y = count))+labs(
    title = "The number of active authors over the hours",
       subtitle = "(12:00 - 24:00)",
       x = "Hour",
       y = "Activity volume"
    ) + 
  geom_col()+theme_classic()+geom_smooth()+changeFont()
```
```{r vis-by-yr}

ggplot(data = grp_yr)+geom_line(aes(year,Tone,colour  = "Tone"))+

  geom_line(aes(year,Authentic,colour  = "Authentic"))+

  geom_line(aes(year,Clout, colour  = "Clout"))+

  geom_line(aes(year,Analytic,colour ="Analytic"))+
  scale_colour_manual("",values =
                        c( "Tone"="blue", 
                           "Authentic"="green", 
                           "Clout" = "black", 
                           "Analytic" = "red"))+ 
  ylim(0.36,.64)+labs(
    title = "The trend of summary variables over the years",
       subtitle = "(2002-2011)",
       x = "Year",
       y = "Active Authors"
    )+theme_minimal()+changeFont()
```

Since Clout's change has been most turbulent, it is worth know which variables define this variable. 
We will use linear regression model to know it 
      --------------------------------------------------------------------------------------------
      2 Is there a relationship between variables?
      --------------------------------------------------------------------------------------------
      

```{r}
library(corrgram)
 
corrgram(data_tidy_norm,upper.panel=panel.cor, main= "The correlation between linguistic variables")

```
   
      
      
      
==============================================================================================================
- Q2

    Analyse the language used by groups. Some starting points:

    a Threads indicate groups of participants communicating on the same topic. Describe the
    threads present in your data.
    --------------------------------------------------------------------------------------------

```{r}
df_tone <- data_tidy %>% group_by(ThreadID) %>% summarise(Tone = median(Tone)) 

summary(df_tone$Tone)
```


    
```{r day-neg-pos}
df <- data_tidy %>% 
  group_by(ThreadID) %>%
  summarise(pos = sum(posemo, na.rm = TRUE),
            neg = sum(negemo, na.rm = TRUE),
            anxiety = sum(anx, na.rm =TRUE))
```
      


```{r day}
df_2 <- data_tidy %>% 
  group_by(ThreadID) %>%
  summarise(Tone = median(Tone, na.rm = TRUE))
df_2 = df_2 %>% mutate(emo = ifelse(Tone >50 , "Positive", "Negative"))%>% 
  group_by(emo) %>% summarise(count =n())

df_2 <- df_2 %>% mutate(pct = prop.table(count))

     ggplot(df_2,aes(x= emo, y =count, label = scales::percent(pct))) + 
    geom_col(position = 'dodge',width = .5) + 
    geom_text(position = position_dodge(width = .9),    # move to center of bars
              vjust = -0.5,    # nudge above top of bar
              size = 3) + coord_flip()
     
     
     
df_for_donut_chart <- df_2 


# Compute percentages
df_for_donut_chart$fraction <- df_for_donut_chart$count / sum(df_for_donut_chart$count)

# Compute the cumulative percentages (top of each rectangle)
df_for_donut_chart$ymax <- cumsum(df_for_donut_chart$fraction)

# Compute the bottom of each rectangle
df_for_donut_chart$ymin <- c(0, head(df_for_donut_chart$ymax, n=-1))

# Compute label position
df_for_donut_chart$labelPosition <- (df_for_donut_chart$ymax + df_for_donut_chart$ymin) / 2

# Compute a good label
df_for_donut_chart$label <- paste0(df_for_donut_chart$category, "\n Percentage: ", round(df_for_donut_chart$pct*100,digits = 2) ,"%")

# Make the plot
ggplot(df_for_donut_chart, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=emo)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=6) +
  scale_fill_brewer(palette=3) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none")

```
    1. Sentiments : are most the thread +ve ?
    2. Pronoun : What are the most used pronoun?
    3. Structure : What are the mean of WC and WPS?



    b By analysing the linguistic variables for all or some of the threads, is it possible to see a
    difference in the language used by different groups?
    --------------------------------------------------------------------------------------------
    
    Find out the languages used between the most postivie and negative threads.
        1. find out the 10 most active threads
    
```{r top-10-Active}
data_tidy_norm <- data_tidy_norm %>% cbind(ThreadID = data_tidy$ThreadID)
df_active_10 <-  data_tidy %>% group_by(ThreadID)%>% summarise(count = n()) %>% arrange(desc(count))
df_active_10 <- head(df_active_10,10)
df_active_10

df <-  data_tidy %>% filter(data_tidy_norm$ThreadID %in% df_active_10$ThreadID  ) %>% arrange(ThreadID)
```
      2. In the most active threads, find out the most most postivie and negative threads. The data set is called data_nega_pose

```{r which-+ve}
data_nega_pose <-  df %>% group_by(ThreadID) %>% summarise(mean = mean(Tone))
data_nega_pose1 <- data_nega_pose
data_nega_pose = data_nega_pose %>% mutate(emo = ifelse(mean >50 , "Positive", "Negative"))%>% 
  group_by(emo) %>% arrange(desc(mean))


#It is noticed that Thread 472752 is the most Positive and 309286 is the most Negative
#most positve
data.mostPose <- data_tidy %>% filter(ThreadID =="472752" )
data.mostPose <- data.mostPose[5:19]
data.mostPose1 <- data.mostPose

#most negative
data.mostNega <- data_tidy %>% filter(ThreadID =="309286" )
data.mostNega <- data.mostNega[5:19]
data.mostNega1 <- data.mostNega
```


Based on the data collected from the most negative and positive, we can make a Radarchart to visualise the difference
```{r radar-chart-all-new }
library(fmsb)
data.mostPose <- data.mostPose %>% select(Analytic,Clout,Authentic,WC,WPS,affect)


data.mostPose <- as.data.frame(lapply(data.mostPose[,], min_max_normalisation))
data.mostPose <- data.mostPose %>% summarise(Analytic_m = mean(Analytic)*100,
                                             Clout_m = mean(Clout)*100,
                                             Authentic_m = mean(Authentic)*100,
                                            WC_m = mean(WC)*100,
                                            WPS_m = mean(WPS)*100,
                                            affect_m = mean(affect)*100
                                             )

# negative
data.mostNega <- data.mostNega %>% select(Analytic,Clout,Authentic,WC,WPS,affect)


data.mostNega <- as.data.frame(lapply(data.mostNega[,], min_max_normalisation))
data.mostNega <- data.mostNega %>% summarise(Analytic_m = mean(Analytic)*100,
                                             Clout_m = mean(Clout)*100,
                                             Authentic_m = mean(Authentic)*100,
                                            WC_m = mean(WC)*100,
                                            WPS_m = mean(WPS)*100,
                                            affect_m = mean(affect)*100
                                             )

radar_data_Pose <- data.mostNega %>% rbind(data.mostPose )



radar_data_Pose <-data.frame( Analytic = c(75, 0 , data.mostPose[1,1],data.mostNega[1,1] ),
                              Clout = c(75, 0 , data.mostPose[1,2],data.mostNega[1,2]  ),
                              Authentic = c(75, 0 , data.mostPose[1,3],data.mostNega[1,3]  ),
                              WC = c(75, 0 , data.mostPose[1,4],data.mostNega[1,4]  ),
                              WPS = c(75, 0 , data.mostPose[1,5],data.mostNega[1,5]  ),
                              affect = c(75, 0 , data.mostPose[1,6],data.mostNega[1,6]  ),
                              row.names = c("max","min","Positive","Negative")
                              )
#defien the colors filled
colors_fill <- c(scales::alpha("yellow", 0.3),scales::alpha("black", 0.5))
#define the line colors
colors_line <- c(scales::alpha("black", 0.5),scales::alpha("darkgrey", 0.5))

 radarchart(radar_data_Pose, axistype = 1,
           seg = 2,
    # Customize the polygon
     pfcol =colors_fill, plwd = 2, plty = 1,
    # Customize the grid
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    pcol = colors_line,
    # Customize the axis
    axislabcol = "grey",
    caxislabels = c(25, 50, 75),
    palcex = 1.5) 

  title(main = "The linguistic variables comparsion between the most negative \nand positive threads",
      cex.main = 1.1,   
      font.main= 1,
      cex.sub = 0.75, font.sub = 1, col.sub = "green",
      col.lab ="darkblue")
      
           
legend( x=1.3,y=1.3,legend = row.names( radar_data_Pose[3:4,] ) ,
        bty = "n",
        pch = 20,
        col=colors_fill,
        cex=1,
        pt.cex=3)

```

```{r radar-chart-all-pronoun }
library(fmsb)
data.mostPose1 <- data.mostPose1 %>% select(i ,we,they, you)
data.mostPose1 <- as.data.frame(lapply(data.mostPose1[,], min_max_normalisation))
data.mostPose1 <- data.mostPose1 %>% summarise(i_m = mean(i)*100,
                                             we_m = mean(we)*100,
                                             you_m = mean(you)*100,
                                            they_m = mean(they)*100
                                             )

# negative
data.mostNega1 <- data.mostNega1 %>% select(i ,we,they, you)
data.mostNega1 <- as.data.frame(lapply(data.mostNega1[,], min_max_normalisation))
data.mostNega1 <- data.mostNega1 %>% summarise(i_m = mean(i)*100,
                                             we_m = mean(we)*100,
                                             you_m = mean(you)*100,
                                            they_m = mean(they)*100
                                             )


radar_data_Pose <- data.mostNega1 %>% rbind(data.mostPose1 )



radar_data_Pose <-data.frame( i = c(30, 0 , data.mostPose1[1,1],data.mostNega1[1,1] ),
                              we = c(30, 0 , data.mostPose1[1,2],data.mostNega1[1,2]  ),
                              you = c(30, 0 , data.mostPose1[1,3],data.mostNega1[1,3]  ),
                              they= c(30, 0 , data.mostPose1[1,4],data.mostNega1[1,4]  ),
                              row.names = c("max","min","Positive","Negative")
                              )
#defien the colors filled
colors_fill <- c(scales::alpha("yellow", 0.3),scales::alpha("black", 0.5))
#define the line colors
colors_line <- c(scales::alpha("black", 0.5),scales::alpha("darkgrey", 0.5))

 radarchart(radar_data_Pose, axistype = 1,
           seg = 2,
    # Customize the polygon
     pfcol =colors_fill, plwd = 2, plty = 1,
    # Customize the grid
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    pcol = colors_line,
    # Customize the axis
    axislabcol = "grey",
    caxislabels = c(10, 20, 30)) 

  title(main = "The pronoun variables comparsion between \nthe most negative and positive threads",
      cex.main = 1.1,   
        font.main= 1,
      cex.sub = 0.75, font.sub = 1, col.sub = "green",
      col.lab ="darkblue")
      
           
legend( x=1.3,y=1.3,legend = row.names( radar_data_Pose[3:4,] ) ,
        bty = "n",
        pch = 20,
        col=colors_fill,
        cex=1,
        pt.cex=3)
```

What about the pronoun they use?
      
    
    c Does the language used within threads (or between threads) change over time? How
        consistent or variable is the language used within threads?
    --------------------------------------------------------------------------------------------
        
        
      C.1  Does the language used within threads (or between threads) change over time? 
        

        
        
        
        
        