---
title: "Week 3 - Introduction to Data Analysis"
author: "ETC1010-5510"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

## <span style="color:blue"> Exercise 2: Exploring Pedestrian Counts </span>

>  Open data-viz2.Rmd

> In here, we'll explore trends in pedestrian counts for Melbourne Central station.
To solve this exercise, fill in the blanks and then knit the final document.

### Load packages and get data
```{r package, message = FALSE, warning = FALSE}
library(tidyverse)
library(rwalkr)
library(lubridate)
```

### Filter out sensor data for Melbourne Central and store in a variable called `walk`

```{r rwalkr-load}
walk_all <- melb_walk_fast(year = 2019)

walk <- walk_all %>% filter(Sensor == "Melbourne Central")

# write_csv(walk, path = "exercises/3b/data/walk_2019.csv")
# write_csv(walk, path = "data/walk_2019.csv")
```

### Read the data from "data" folder and convert to "tidy" format

```{r read-walk}
walk <- read_csv("data/walk_2019.csv")
walk
```

### Conversion to Tidy data: Add new variables to data: `month` and `wday`
```{r month-week}
walk_tidy <- walk %>%
  mutate(month = month(Date,  label = TRUE, abbr = TRUE), 
         wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1))
walk_tidy
```

## Answer the following questions:

### Q1. Generate a plot for _monthly_ pedestrian counts. Use `month` and `Count`.

```{r month, eval = T}
ggplot(walk_tidy,
       aes(x = month, 
           y = Count)) + 
  geom_col()
```

### Q2. Similarly, generate a plot for _weekly_ pedestrian counts.

```{r week, eval = T}
ggplot(walk_tidy,
       aes(x = wday, 
           y = Count)) + 
  geom_col()
```

### Q3. Create above graphs (_monthly, weekly and daily_ counts) for Birrarung Marr.

> First, let's pull the 2019 data for the sensor Birrarung Marr. Transform data into tidy form.

> Hint: Look at the similar code for Melbourne Central above.

```{r walk-burra}
# filter data for Birrarung Marr
walk_birra <- walk_all %>% filter(Sensor == "Birrarung Marr")

# transform to tidy data; create new columns: `month` and `wday`
walk_birra_tidy <- walk_birra %>%
  mutate(month = month(Date,  label = TRUE, abbr = TRUE), 
         wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1))
walk_birra_tidy
```
> Next, create plots:

```{r plots-birra}

# Monthly counts:
ggplot(walk_birra_tidy,
       aes(x = month, 
           y = Count)) + 
  geom_col()

# Weekly counts:
ggplot(walk_birra_tidy,
       aes(x = wday, 
           y = Count)) + 
  geom_col()
```

> Bar plots do not provide sufficient insights:
>
>* There might be a different number of days of the week over the year.
>* This means that simply summing the counts might lead to a misinterpretation of pedestrian patterns.
>* Similarly, months have different numbers of days.

### Q4. Generate a line-plot for _daily_ pedestrian counts. Here, we'll use `geom_line()` and `group` aesthetic to group the lines.

> Hint: Use `Time` and `Count` as aesthetics.

```{r daily, out.width=c('50%', '50%'), fig.show='hold', eval = T}
# Daily counts- Melbourne central
ggplot(walk_tidy,
       aes(x = Time, 
           y = Count,
           group = Date)) + 
  geom_line() +
  ggtitle("Daily counts: Melbourne Central")

ggplot(walk_birra_tidy,
       aes(x = Time, 
           y = Count,
           group = Date)) + 
  geom_line() +
  ggtitle("Daily counts: Birrarung Marr")
```

>Line charts reveal interesting patterns from daily pedestrian counts.

### **Is there the same _daily_ pattern as observed at Melbourne Central? If not, how does it differ?**
> It is hard to tell because of different y-scales. The y-scales need to be the same to make a valid comparison.

### Q5. What are the number of pedestrians per day?
> Hint: Refer to Workshop slides

```{r count-day}
walk_day <- walk_tidy %>% 
  group_by(Date) %>%
  summarise(day_count = sum(Count, na.rm = TRUE))
walk_day
```

### Q6. What are the average number of pedestrians per weekday?
> Hint: Refer to Workshop slides

```{r count-weeekday}
walk_day %>%
  mutate(wday = wday(Date, label = TRUE, abbr = TRUE, week_start = 1)) %>% 
  group_by(wday) %>%
  summarise(m = mean(day_count, na.rm = TRUE),
            s = sd(day_count, na.rm = TRUE))
```

## <span style="color:blue"> Exercise 3: More Data Visualisation using Calendar plots</span>

> We'll continue in this same .Rmd

> Let's combine the Birrarung Marr data with Melbourne central data (walk) using `bind_rows()`. Or even subset the original data that has all the obsercations into something that has melbourne central AND birrarung marr. There are many ways to get to the same answer.

> Let's bind the two Sensors- Melbourne central and Birrarung Marr together

```{r bind-both}
walk_both <- bind_rows(walk_tidy,
                       walk_birra_tidy)
```

## Answer the following questions:

### Q1. Is there the same _daily_ pattern as observed at Melbourne Central? If not, how does it differ? Can you provide reason for the different daily pattern?

```{r walk-both}
ggplot(walk_both,
        aes(x = Time, 
            y = Count,
           colour = Sensor,
           group= Date)) + 
     geom_line() +
  facet_wrap(~Sensor)

```

> No. While pedestrian counts at Melbourne Central follow a strict pattern, Birrarung Marr has a lot of variation. 

> This could be because of the varied nature of activities at the two Sensor locations. Melbourne Central hosts activities such as office commutes, student's commute, shopping, restaurants etc.

> Birrarung Marr Park on the other hand hosts events and festivals such as Moomba fest! The park also provides walking, cycle access and other recreational activities.

### Q2. Can you detect special events, like the Moomba Festival, based on pedestrian counts?
> We can do this using Calendar plots. Use your Google skills to find out Moomba Fest 2019 dates and observe pedestrian trends.

```{r cal}
library(sugrrants)
walk_tidy_calendar <- 
  frame_calendar(walk_tidy,
                 x = Time, 
                 y = Count, 
                 date = Date, 
                 nrow = 4)
p1 <- ggplot(walk_tidy_calendar,
       aes(x = .Time, 
           y = .Count, 
           group = Date)) +
  geom_line()
prettify(p1)
```

> Also, look at the pedestrian counts on 31st Dec. What do you observe?

### Q3. How do the calendar plots compare for Melbourne Central and Birrarung Marr?
```{r cal-both}
walk_both_tidy_calendar <- 
  frame_calendar(walk_both,
                 x = Time, 
                 y = Count, 
                 date = Date, 
                 nrow = 4)

p2 <- ggplot(walk_both_tidy_calendar,
       aes(x = .Time, 
           y = .Count,
           colour= Sensor,
           group = Date)) +
  geom_line() +
  facet_wrap(~Sensor)
prettify(p2)
```

### Q4. Join holidays data with pedestrian counts and observe trends. Can you match these with known events, found from web searching?

```{r holidays}
library(tsibble)
library(sugrrants)
library(timeDate)
vic_holidays <- holiday_aus(2019, state = "VIC")
vic_holidays

walk_holiday <- walk_tidy %>%
  mutate(holiday = if_else(condition = Date %in% vic_holidays$date, 
                          true = "yes", 
                          false = "no")) %>%
  mutate(holiday = if_else(condition = wday %in% c("Sat", "Sun"), 
                          true = "yes", 
                          false = holiday)) 
walk_holiday

walk_holiday_calendar <- frame_calendar(data = walk_holiday,
                                        x = Time, 
                                        y = Count, 
                                        date = Date, 
                                        nrow = 6)

p3 <- ggplot(walk_holiday_calendar,
       aes(x = .Time, 
           y = .Count, 
           group = Date, 
           colour = holiday)) +
  geom_line() + 
  scale_colour_brewer(palette = "Dark2")
prettify(p3)
```

> Now, go to `knitr::opts_chunk` at line 9, change `eval = TRUE`, then knit the document.

### What to do if you are stuck?

* Read through the lecture slides again, they most likely have an answer to these questions!
* Take the question and copy-paste it into google, adding "lubridate" to the google query.

