---
title: "Week 3 - Introduction to Data Analysis"
author: "ETC1010-5510"
output:
  html_document: default
---

# <span style="color:blue"> Exercise 1: Data Visualisation and Lubridate </span>
```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

### Load packages

```{r load-packages, warning = FALSE, message = F}
library(readr)
library(rwalkr)
library(lubridate)
library(tidyverse)
```

## Answer the following questions:

### <span style="color:green"> Section A: Play with dates </span> 

### Convert something of the format "2020-01-20" into a date?
```{r add-tz-ymd}
ymd("2020-01-20")
```

### How do you show today's date?
```{r show-today}
now()
```

### Q1. Convert something of the format "2020-01-20 10:00:01" into a date time? Try function: `ymd_hms()`

```{r ymd-hms}
a <- c("2020-01-20 10:00:01")
b <- ___("2020-01-20 10:00:01")
```

> Compare a and b. Try `class(a)` and `class(b)`.

### Q2. Get the weekday and name of weekday out of "2020-01-20". Try `wday()`
```{r wday-and-options}
wday("2020-01-20")
___("2020-01-20", label=T)
```

### Q3. Get the month and month name out of "2020-01-20". Try `label`

```{r month-label}
month("2020-01-20")
___("2020-01-20", ___)
```

### Q4. Get the un-abbreviated month name out of "2020-01-20"

```{r month-label-abbr}
___("2020-01-20", ___, abbr  = FALSE)
```

### Q5. Get the day of the year out of "2020-01-20". Try `yday()`

```{r yday}
___("2020-01-20")
```

### <span style="color:green"> Section B: Data Visualisation </span> 

### Load data

```{r read-data-in, warning = FALSE}
walk_tidy <- read_csv(file = "data/walk-feb-march-2020.csv.gz")
```

### Q1. Make a side-by-side boxplot of pedestrian counts by sensor. Use `Count` and `Sensor` in aesthetics. 

>Make sure to load the ggplot2 library

```{r gg-walk-point}
library(ggplot2)
ggplot(walk_tidy,
       aes(x = ___,
           y = ___)) +
  geom_point()
```

>There are many observations presented with `geom_point()`!

### Q2. Try using `geom_boxplot()` instead to summarise this information.

```{r gg-walk-boxplot}
ggplot(walk_tidy,
       aes(x = ___,
           y = ___)) +
  ____
```

>Describe what you learn from the boxplot.

### Q3. Subset to Monday of the first week of March using `filter`

> Hint: read the previous description of the `dmy` function to work out how to choose the date.

```{r create-first-march}
walk_first_march <- walk_tidy %>%
  ___(Date >= ___("01-03-2020"), 
         Date <= ___("08-03-2020"))
```

### Q4. Using the first week data, make line plots of `Count` by `Date_Time`, separately for each sensor. 

```{r gg-first-week}
ggplot(walk_first_march,
       aes(x = ___,
           y = ___)) +
  geom_line() +
  facet_wrap(~Sensor,
             ncol = 1)
```

> What do you learn? Explain the patterns.

> Now, go to `knitr::opts_chunk` at line 9, change `eval = TRUE`, then knit the document.

# Appendix: Code using which the data was obtained.

## Reading in the data

Pull a copy of the Melbourne pedestrian sensor data for March, 2020, using the `rwalkr` package, filter to keep two sensors "Melbourne Central", "Flinders Street Station Underpass". You can use the code below, or read in the data from the "data" folder (it might take a few minutes to download the data - it took me about 1 minute on a 2016 Macbook Pro with 16Gb RAM). 

**(Note that the code chunk below is set to `eval = FALSE`, which means that it will not run when you knit the document.)**

### Explanation of reading / tidying code

* First we use `library(rwalkr)` to load up the libraries that are needed to download the pedestrian tracking data - this is the `rwalkr` R package
* Then we load the `lubridate` R package with `library(lubridate)`. Lubridate is an R package that we will cover in more detail in the next lecture. For the moment all you need to know about is the `dmy()` function, which takes a character input (like "20-03-2020") and converts it into a special date format.
* You can explore what this format is by looking at the `start_date` and `end_date` objects. 
* Here we are using the foramt: "ddmmyyy", where there are two digits for the day, two digits for the month and four for the year.
* So, "01012000" would be the First of January, 2000.

```{r read-data, eval = FALSE}
  library(rwalkr)
  library(lubridate)
  library(tidyverse)

 start_date <- dmy("01-02-2020")
 end_date <- dmy("30-03-2020")

 walk_raw <- melb_walk(from = start_date,
                       to = end_date)
```

There are many sensors that are measured by the pedestrian tracking system. You can find out more information about the data by looking up the helpfil for the `melb_walk` function by typing `?melb_walk`.

We want to `filter` the data based on the sensor, and only look at two: Melbourne Central, and the Flinders Street Station Underpass. We need to make sure we have loaded the `dplyr` library

```{r filter-data, eval = FALSE}
 library(dplyr)
 walk_tidy <- walk_raw %>%
   filter(Sensor %in% c("Melbourne Central",
                        "Flinders Street Station Underpass"))
```

Next we save the data into a (compressed) CSV format. You can run this code chunk below ("read-data") if you want to read the data from the "data" folder. **Remember that if you knit this document, you will need to set the code chunk from `` to `eval = TRUE`.**

```{r read-data-2, eval = FALSE}
 library(readr)
 write_csv(walk_tidy,
           path = "data/walk-feb-march-2020.csv.gz")

 write_csv(walk_raw,
           path = "data/walk_raw.csv.gz")

```
