```{r}
library(ggplot2)
set.seed(9999) # Random seed to make subset reproducible
dsmall <-diamonds[sample(nrow(diamonds), 1000), ] # sample of 1000 rows
```

```{r}

lm <-  lm(data= dsmall,price~carat )
summary(lm)
```


```{r}
attach(dsmall)

contrasts(clarity) = contr.treatment(8)
contrasts(clarity)
```

color
```{r}
attach(dsmall)

contrasts(color) = contr.treatment(7)
contrasts(color)
```
```{r}
attach(dsmall)

contrasts(cut) = contr.treatment(5)
contrasts(cut)
```

```{r}
#r-square = how much the variation of X can affect Y
#when null hypo is rejected, that means the coffient is statistically siginificance
car clat colo cut
lm (log(price)~ log(carat)+clarity+color+cut,data= dsmall)
```



Q2
```{r}
B = read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Week-(7)/body.dat.csv")
attach(B)
# Test the hypothesis that men are taller than women on average.
#Assume a significance of 5% 

t.test(Height[Gender == "Male"],
       Height[Gender == "Female"], 
       conf.level = .95)

##Conclusion :
##if p v is less than 5%, so we can accept the alternate hypo
## at 95%,men are taller than women on average.

```

```{r}
B = read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Week-(7)/body.dat.csv")
attach(B)
# Test the hypothesis that men are taller than women on average.
#Assume a significance of 5% 

t.test(Height[Gender == "Male"],
       Height[Gender == "Female"], 
       conf.level = .99)

##Conclusion :
##if p v is less than 1%, so we can accept the alternate hypo
## at 99%,men are taller than women on average.
```
Q3 : look at the foto
```{r}
B$BMI = B$Weight / (B$)


```
 (d) Calculate the regression of Height on the other body measurements for men and 
 women separately. Which measurements are the most significant predictors of height 
 for each gender? 

 
```{r}
attach(B)
B.Male = B[which(B$Gender == "Male")]
B.Female = B[which(B$Gender == "Female")]

male.fit = lm(Height ~ .)
```

 

3. The data file “Dunnhumby1-20.csv” is a cut down and modified set of test data from the 
Kaggle competition to predict when consumers would next visit a Dunnhumby supermarket 
and how much they would spend. See: http://www.kaggle.com/c/dunnhumbychallenge for 
more information. The current modified data set contains the customer ID, Date of visit, 
Days since last visit (Delta), and Spend for 20 customers from the test set. 

 

Calculate the regression of Spend vs Delta for each customer and summarize the results in a data frame similar to that below. 
Hint: try using “plyr” package and dlply function. 

 

```{r}
library("plyr")
c <-read.csv("C:/Users/sjsa3/Desktop/Shared_with_Mac/year2_sem1/FIT3152/Week-(7)/Dunnhumby1-20.csv")
attach(c)
  models = dlply(c,"customer_id",function(df) lm( visit_spend ~ visit_delta,data= df))

  
  d = ldply(models, coef)
  View(d)
```



4. Using the data from the UCI Machine Learning Repository comment on the factors affecting 
red wine quality. Data site is: http://archive.ics.uci.edu/ml/datasets/Wine+Quality The file name 
is: winequality-red.csv. 

```{r}
rwine = read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", sep = ";")

View(rwine)
#comment on the factors affecting red wine quality.

fit = lm(data= rwine, quality ~ .)
summary(fit)
```
 5. Install the “ISLR” library. Using the “Carseats” data, calculate the regression equation 
predicting Sales (child car seat sales) as a function of the input variables. Which variables 
are significant predictors? 
```{r}
library(ISLR)
C = Carseats
str(C)
```

```{r}
attach(C)
contrasts(ShelveLoc)   = contr.treatment(3)
contrasts(ShelveLoc)
```

```{r}
contrasts(US)   = contr.treatment(2)
contrasts(US) 

contrasts(Urban)   = contr.treatment(2)
contrasts(Urban)

Cmodel = lm(Sales ~ . ,data=C)
summary(Cmodel)
```

6. The text, G. James et al., An Introduction to Statistical Learning: with Applications in R 
(ISLR) uses the “Advertising” data set to illustrate a number of different learning models. A 
description of the data (p15) follows: “The Advertising data set consists of the sales of that 
product in 200 different markets, along with advertising budgets for the product in each of 
those markets for three different media: TV, radio, and newspaper. A copy of the data was 
downloaded from: https://www.kaggle.com/ashydv/advertising-dataset and is on Moodle. 

 

 Using the Advertising data, answer the following questions (taken from pp59-60 ISLR): 

 

(a) Is there a relationship between advertising budget and sales? 
```{r}
fit = lm(data=C, Sales ~ Advertising)

summary(fit)
```

(b) How strong is this relationship? 
(c) Is the relationship linear? 
(d) Which media contribute to sales? 
(e) How accurately can we estimate the effect of each medium on sales? 
(f) (Extension) Is there synergy (interactions) among the advertising media? 

 

 Potential ways of addressing these questions using regression models and extensive 
discussion of regression can be found on pages 59-82 of ISLR. 
```{r}

```

